### 2017-04-19 presentations/2017-04-19_itamp/present.tex
[presentations/2017-04-19_itamp] Tweaks after practice talk

\documentclass{beamer}
\mode<presentation>{
  \usetheme{Boadilla}
  \usefonttheme[onlylarge]{structurebold}
  \usefonttheme[stillsansseriflarge]{serif}
  \setbeamerfont*{frametitle}{size=\normalsize,series=\bfseries}
  % \setbeamertemplate{navigation symbols}{}
  \setbeamercovered{transparent}
}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{esint}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{xkeyval}
\usepackage{xargs}
\usepackage{verbatim}
\usepackage{listings}
\usepackage{multimedia}
\usepackage{bm}
\usepackage{siunitx}
\usetikzlibrary{
  arrows,
  calc,
  decorations.pathmorphing,
  decorations.pathreplacing,
  decorations.markings,
  fadings,
  positioning,
  shapes
}

\mode<handout>{
  \usepackage{pgfpages}
  \pgfpagesuselayout{4 on 1}[a4paper,landscape,border shrink=5mm]
  \setbeamercolor{background canvas}{bg=black!10}
}

\newcommand\pgfmathsinandcos[3]{%
  \pgfmathsetmacro#1{sin(#3)}%
  \pgfmathsetmacro#2{cos(#3)}%
}
\newcommand\LongitudePlane[3][current plane]{%
  \pgfmathsinandcos\sinEl\cosEl{#2} % elevation
  \pgfmathsinandcos\sint\cost{#3} % azimuth
  \tikzset{#1/.estyle={cm={\cost,\sint*\sinEl,0,\cosEl,(0,0)}}}
}
\newcommand\LatitudePlane[3][current plane]{%
  \pgfmathsinandcos\sinEl\cosEl{#2} % elevation
  \pgfmathsinandcos\sint\cost{#3} % latitude
  \pgfmathsetmacro\yshift{\cosEl*\sint}
  \tikzset{#1/.estyle={cm={\cost,0,0,\cost*\sinEl,(0,\yshift)}}} %
}
\newcommand\DrawLongitudeCircle[2][1]{
  \LongitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{atan(sin(#2)*cos(\angEl)/sin(\angEl))} %
  \draw[current plane] (\angVis:1) arc (\angVis:\angVis+180:1);
  \draw[current plane,dashed] (\angVis-180:1) arc (\angVis-180:\angVis:1);
}
\newcommand\DrawLatitudeCircleArrow[2][1]{
  \LatitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  \pgfmathsetmacro\sinVis{sin(#2)/cos(#2)*sin(\angEl)/cos(\angEl)}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{asin(min(1,max(\sinVis,-1)))}
  \draw[current plane,decoration={markings, mark=at position 0.6 with {\arrow{<}}},postaction={decorate},line width=.6mm] (\angVis:1) arc (\angVis:-\angVis-180:1);
  \draw[current plane,dashed,line width=.6mm] (180-\angVis:1) arc (180-\angVis:\angVis:1);
}
\newcommand\DrawLatitudeCircle[2][1]{
  \LatitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  \pgfmathsetmacro\sinVis{sin(#2)/cos(#2)*sin(\angEl)/cos(\angEl)}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{asin(min(1,max(\sinVis,-1)))}
  \draw[current plane] (\angVis:1) arc (\angVis:-\angVis-180:1);
  \draw[current plane,dashed] (180-\angVis:1) arc (180-\angVis:\angVis:1);
}
\newcommand\coil[1]{
  {\rh * cos(\t * pi r)}, {\apart * (2 * #1 + \t) + \rv * sin(\t * pi r)}
}
\makeatletter
\define@key{DrawFromCenter}{style}[{->}]{
  \tikzset{DrawFromCenterPlane/.style={#1}}
}
\define@key{DrawFromCenter}{r}[1]{
  \def\@R{#1}
}
\define@key{DrawFromCenter}{center}[(0, 0)]{
  \def\@Center{#1}
}
\define@key{DrawFromCenter}{theta}[0]{
  \def\@Theta{#1}
}
\define@key{DrawFromCenter}{phi}[0]{
  \def\@Phi{#1}
}
\presetkeys{DrawFromCenter}{style, r, center, theta, phi}{}
\newcommand*\DrawFromCenter[1][]{
  \setkeys{DrawFromCenter}{#1}{
    \pgfmathsinandcos\sint\cost{\@Theta}
    \pgfmathsinandcos\sinp\cosp{\@Phi}
    \pgfmathsinandcos\sinA\cosA{\angEl}
    \pgfmathsetmacro\DX{\@R*\cost*\cosp}
    \pgfmathsetmacro\DY{\@R*(\cost*\sinp*\sinA+\sint*\cosA)}
    \draw[DrawFromCenterPlane] \@Center -- ++(\DX, \DY);
  }
}
\newcommand*\DrawFromCenterText[2][]{
  \setkeys{DrawFromCenter}{#1}{
    \pgfmathsinandcos\sint\cost{\@Theta}
    \pgfmathsinandcos\sinp\cosp{\@Phi}
    \pgfmathsinandcos\sinA\cosA{\angEl}
    \pgfmathsetmacro\DX{\@R*\cost*\cosp}
    \pgfmathsetmacro\DY{\@R*(\cost*\sinp*\sinA+\sint*\cosA)}
    \draw[DrawFromCenterPlane] \@Center -- ++(\DX, \DY) node {#2};
  }
}
\makeatother

% not mandatory, but I though it was better to set it blank
\setbeamertemplate{headline}{}
\def\beamer@entrycode{\vspace{-\headheight}}

\tikzstyle{snakearrow} = [decorate, decoration={pre length=0.2cm,
  post length=0.2cm, snake, amplitude=.4mm,
  segment length=2mm},thick, ->]

%% document-wide tikz options and styles

\tikzset{%
  % >=latex, % option for nice arrows
  inner sep=0pt,%
  outer sep=2pt,%
  mark coordinate/.style={inner sep=0pt,outer sep=0pt,minimum size=3pt,
    fill=black,circle}%
}
\tikzset{
  % Define standard arrow tip
  >=stealth',
  % Define style for boxes
  punkt/.style={
    rectangle,
    rounded corners,
    draw=black, very thick,
    text width=8em,
    minimum height=2.5em,
    text centered},
}
\makeatletter
\newbox\@backgroundblock
\newenvironment{backgroundblock}[2]{%
  \global\setbox\@backgroundblock=\vbox\bgroup%
  \unvbox\@backgroundblock%
  \vbox to0pt\bgroup\vskip#2\hbox to0pt\bgroup\hskip#1\relax%
}{\egroup\egroup\egroup}
\addtobeamertemplate{background}{\box\@backgroundblock}{}
\makeatother

% \def\timeleft{15:00->14:55}

\title[Na ground state cooling]{Raman sideband cooling of a single sodium atom to 3D ground state}
\date{April 19, 2017}
\author[Yichao Yu]{Yichao Yu\\
  \vspace{0.5cm}
  {\footnotesize Lee Liu, Dr. Nick Hutzler,\\
    Jessie Zhang, Dr. Jon Hood}}
\institute{Ni Group/Harvard}

\begin{document}

\begin{frame}{}
  \titlepage
\end{frame}

% \begin{frame}{}
%   \tableofcontents
% \end{frame}

\begin{frame}{}
  \begin{center}
    \begin{tikzpicture}
      \visible<1> {
        \path (0, 0) node
        {\includegraphics[width=10cm]{../2015-05-05_cua/overall_full_min.png}};
      }
      \visible<2-> {
        \path (0, 0) node
        {\includegraphics[width=10cm]{../2015-05-05_cua/overall_trap_min.png}};
        \path (3, -3) node
        {\includegraphics[width=3cm]{../../experiments/nacs_atoms/imgs/single_viridis.png}};
      }
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}{Wave function size mismatch}
  \begin{center}
    \begin{tikzpicture}
      \shade[ball color=blue!90] (0, 0) circle (0.45);
      \shade[ball color=orange!90] (0.4, 0.2) circle (0.3);
      \draw[<->,line width=1] (-0.45, 0.4) -- (0.35, 0.8);
      \path (-0.05, 0.6) node[rotate=26.565,above=2pt] {$4a_0$};
      \path (0.2, -2) node[below] {\textbf{Molecule}};

      \draw[line width=1] plot[samples=200,domain=-2:2,variable=\x] ({\x + 6}, {(\x)^2 * 0.8 - 1.5});
      \draw[line width=2,orange!90!black]
      plot[samples=200,domain=-2:2,variable=\x] ({\x + 6}, {exp(-(\x)^2 * 1.3) - 0.988});
      \draw[line width=0.8] (6 - 0.8, -0.988) -- (6 + 0.8, -0.988);
      \draw[<->,line width=1] (6 - 1, 0.3) -- (6 + 1, 0.3);
      \path (6, 0.3) node[above] {$1000a_0$};
      \path (6, -2) node[below] {\textbf{Atom}};
    \end{tikzpicture}
    \vspace{1cm}
    \begin{block}{Goal of cooling}
      \begin{itemize}
      \item Single initial state
      \item Shrink wavefunction size
      \end{itemize}
    \end{block}
  \end{center}
\end{frame}

\begin{frame}{Raman sideband cooling of Sodium}
  \begin{columns}
    \column{6.5cm}
    \includegraphics[width=6cm]{Na_RSC_schematic.pdf}
    \column{5cm}
    \visible<2->{
      \begin{block}{Difficulties}
        \begin{itemize}
        \item High initial temperature ($40\mu K$)
        \item<3-> High recoil heating\\
          {\footnotesize (High Lamb Dicke parameter)}
        \end{itemize}
      \end{block}
    }
  \end{columns}
\end{frame}

\begin{frame}{Raman sidebands}
  \begin{center}
    \begin{tikzpicture}
      \visible<1> {
        \path (0, 0) node {
          \includegraphics[width=10cm]{../../experiments/misc/imgs/data_spectrum_20170409_r2_before.png}
        };
        \draw[<-, line width=1.5] (-0.7, 0) -- (0, 1.6) node[right] {\Large Carrier};
      }
      \visible<2> {
        \path (0, 0) node {
          \includegraphics[width=10cm]{../../experiments/misc/imgs/data_spectrum_20170409_r2.png}
        };
      }
      \visible<3> {
        \path (0, 0) node {
          \includegraphics[width=10cm]{../../experiments/misc/imgs/data_spectrum_20170409_r3.png}
        };
      }
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}{Raman sidebands}
  \begin{center}
    \begin{tikzpicture}[scale=1.41176]
      \visible<1> {
        \path (0, 0) node {
          \includegraphics[width=12cm]{../../experiments/misc/imgs/data_spectrum_20170409_a1_before.png}
        };
        \draw[red, <-, line width=1.5] (-3.15, 0.5) -- (-2.95, 1.2)
        node[above right] {\Large 1st order heating};
        \draw[<-, line width=1.5] (-2.45, 0.2) -- (-2.25, 0.8) node[above right] {\Large Carrier};
        \draw[black!40!green, <-, line width=1.5] (-1.7, 0.2) -- (-1.55, 0.6)
        node[right] {\Large 1st order cooling};
        \draw[blue, <-, line width=1.5] (-0.95, -0.1) -- (-0.55, 0.1)
        node[right] {\Large 2nd order cooling};
        \path[magenta] (0.25, -0.5) node[right] {\Large And higher orders{$\cdots$}};
      }
      \visible<2-> {
        \path (0, 0) node {
          \includegraphics[width=12cm]{../../experiments/misc/imgs/data_spectrum_20170409_a1.png}
        };
      }
      \visible<3-> {
        \fill[white,opacity=0.95] (-4.2, -2.5) rectangle (5, 2.5);
        \path (0, 1) node[align=center]
        {
          \begin{tabular}{|c|c|}
            \hline
            \textbf{Axis}&\textbf{Ground state probability}\\\hline
            1 (Axial)&93.1(2.5)\%\\\hline
            2 (Radial)&91.9(2.3)\%\\\hline
            3 (Radial)&92.9(2.5)\%\\\hline
          \end{tabular}
        };
        \visible<4-> {
          \path (0, -1) node[align=center]
          {
            \textbf{3D ground state:} $79.5(3.6)\%$\\
            \textbf{Loss after cooling:} $15\%$\\\\
            \textbf{Total 3D ground state preparation fidelity:} $67.6(3.1)\%$
          };
        }
      }
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}{Rabi flopping (radial)}
  \begin{center}
    \visible<+->{
      \includegraphics[width=10cm]{../../experiments/rabi_flop/imgs/fit_20170409_r2_0_ba.png}\\
    }
    \visible<+->{
      Good agreement in ground state probability between spectrum and Rabi flopping data.
    }
  \end{center}
\end{frame}

\begin{frame}{Rabi flopping (axial)}
  \begin{center}
    \includegraphics[height=4.8cm]{../../experiments/rabi_flop/imgs/fit_20170409_a1_0_nol.png}
    \includegraphics[height=4.8cm]{../../experiments/rabi_flop/imgs/fit_20170409_a1_p1_nol.png}
  \end{center}
\end{frame}

\begin{frame}{}
  \begin{block}{Conclusion}
    $67.6(3.1)\%$ ground state preparation fidelity {\small ($79.5(3.6)\%$ without loss)}
  \end{block}
  \vspace{1cm}
  \begin{block}{Improvements}
    \begin{itemize}
    \item Reduce off-resonance scattering from Raman beams
    \item Reduce magnetic field fluctuation
    \item Reduce loss during cooling
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{}
\end{frame}

\begin{frame}{Axial matrix element}
  \begin{center}
    \includegraphics[width=6.5cm]{{../../calculations/sideband_strength/imgs/coupling_0.43_0-6}.png}
  \end{center}
\end{frame}

\begin{frame}{Radial 2 matrix element}
  \begin{center}
    \includegraphics[width=6.5cm]{{../../calculations/sideband_strength/imgs/coupling_0.35_0-2}.png}
  \end{center}
\end{frame}

\begin{frame}{Radial 3 matrix element}
  \begin{center}
    \includegraphics[width=6.5cm]{{../../calculations/sideband_strength/imgs/coupling_0.29_0-2}.png}
  \end{center}
\end{frame}

\end{document}