### 2017-05-22 presentations/na_cool_paper/paper.tex
[presentations/na_cool_paper] WIP

\documentclass[aps,prl,twocolumn,groupedaddress]{revtex4-1}
% \documentclass[aps,twocolumn,secnumarabic,balancelastpage,amsmath,amssymb,nofootinbib]{revtex4-1}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{chapterbib}
\usepackage{color}
\usepackage{graphics}
\usepackage[pdftex]{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{epsf}
\usepackage{bm}
\usepackage{tikz}
\usepackage{asymptote}
\usepackage{thumbpdf}
\usepackage{bbm}
\usepackage{amscd}
\usepackage{units}
\usepackage{textcomp}
\usepackage[utf8x]{inputenc}
\usepackage{feyn}
\usepackage{feynmp}
\usepackage[colorlinks=true]{hyperref}
\newcommand{\drelectron}[1]{\node at #1 [circle, draw, inner sep=0pt, minimum size=1pt] {\_}}

\newcommand{\ud}{\mathrm{d}}
\newcommand{\ue}{\mathrm{e}}
\newcommand{\ui}{\mathrm{i}}
\newcommand{\res}{\mathrm{Res}}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\dsum}{\displaystyle\sum}
\newcommand{\dprod}{\displaystyle\prod}
\newcommand{\dlim}{\displaystyle\lim}
\newcommand{\dint}{\displaystyle\int}
\newcommand{\fsno}[1]{{\!\not\!{#1}}}
\newcommand{\eqar}[1]
{
  \begin{align*}
    #1
  \end{align*}
}
\newcommand{\eqarn}[1]
{
  \begin{align}
    #1
  \end{align}
}
\newcommand{\texp}[2]{\ensuremath{{#1}\times10^{#2}}}
\newcommand{\dexp}[2]{\ensuremath{{#1}\cdot10^{#2}}}
\newcommand{\eval}[2]{{\left.{#1}\right|_{#2}}}
\newcommand{\paren}[1]{{\left({#1}\right)}}
\newcommand{\lparen}[1]{{\left({#1}\right.}}
\newcommand{\rparen}[1]{{\left.{#1}\right)}}
\newcommand{\abs}[1]{{\left|{#1}\right|}}
\newcommand{\sqr}[1]{{\left[{#1}\right]}}
\newcommand{\crly}[1]{{\left\{{#1}\right\}}}
\newcommand{\angl}[1]{{\left\langle{#1}\right\rangle}}
\newcommand{\tpdiff}[4][{}]{{\paren{\frac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}_{#4}}}
\newcommand{\tpsdiff}[4][{}]{{\paren{\frac{\partial^{#1}}{\partial {#3}{}^{#1}}{#2}}_{#4}}}
\newcommand{\pdiff}[3][{}]{{\frac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}}
\newcommand{\diff}[3][{}]{{\frac{\ud^{#1} {#2}}{\ud {#3}{}^{#1}}}}
\newcommand{\psdiff}[3][{}]{{\frac{\partial^{#1}}{\partial {#3}{}^{#1}} {#2}}}
\newcommand{\sdiff}[3][{}]{{\frac{\ud^{#1}}{\ud {#3}{}^{#1}} {#2}}}
\newcommand{\tpddiff}[4][{}]{{\left(\dfrac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}\right)_{#4}}}
\newcommand{\tpsddiff}[4][{}]{{\paren{\dfrac{\partial^{#1}}{\partial {#3}{}^{#1}}{#2}}_{#4}}}
\newcommand{\pddiff}[3][{}]{{\dfrac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}}
\newcommand{\ddiff}[3][{}]{{\dfrac{\ud^{#1} {#2}}{\ud {#3}{}^{#1}}}}
\newcommand{\psddiff}[3][{}]{{\frac{\partial^{#1}}{\partial{}^{#1} {#3}} {#2}}}
\newcommand{\sddiff}[3][{}]{{\frac{\ud^{#1}}{\ud {#3}{}^{#1}} {#2}}}

\begin{document}
\tikzstyle{every picture}+=[remember picture]
\title{Raman sideband cooling of single Na atoms to 3D vibrational ground state}
\author{....}
\date{\today}
\affiliation{Harvard Department of Physics}

\begin{abstract}
  We report Raman sideband cooling of a single neutral Sodium atom to its three-dimensional
  vibrational ground state in an optical tweezer.
  Despite having a very large Lamb-Dicke parameter, high initial temperature and
  large differential AC Stark shift in the excited state,
  after applying a cooling sequence for a hundred milliseconds,
  we observed a ground state preparation fidelity of $70\%$ using sideband thermometry.
  We demonstrated that Raman sideband cooling to vibrational ground state is applicable to
  systems where tight confinement or low initial cooling is hard to achieve.
  For example, the result provides new opportunities to achieve much lower temperatures
  in cold molecules with direct laser cooling.
\end{abstract}

\maketitle

Why do we need to cool (for STIRAP/transfer to molecular state)\\

* Pure initial state\\
* Small wavefunction span\\

\begin{figure*}
  \includegraphics[height=6.5cm]{imgs/Na_RSC_schematic.pdf}
  \includegraphics[height=6.5cm]{sequence.pdf}
  \caption{(A) Energy diagram. D1 OP
    (B) Beam diretions and polarizations
    (C) Sequence description
    \label{f-setup}}
\end{figure*}

We achieve the high ground state preparation fidelity needed for molecule formation
using Raman sideband cooling. The process consists of multiple cycles of laser pulses
to manipulate the internal and vibrational states of the atoms.
Figure \ref{f-setup}A shows the schematic of the energy levels and the cooling sequence
in our setup.
Each cooling cycle starts with the Sodium atom in the $|F=2, m_F=-2\rangle$
ground electronic state and a certain vibration state $n$.
In the first step, a Raman pulse drives a transition to the vibrational state $n-\Delta n$
to reduce the vibrational energy while also changing the internal state to $|F=1, m_F=-1\rangle$.
In the second step, which finishes the cooling cycle,
an optical pumping pulse bring the atom back to the $|F=2, m_F=-2\rangle$ state to take away
the entropy. The second step could also change the vibrational state of the atoms which
could cause heating. The possiblity for this to happen for an atom in the vibrational level $n$
is approximately proportional to the effective Lamb-Dicke parameter
$\eta^{OP}_{eff}=\sqrt{2n+1}\eta^{OP}$ where $\eta^{OP}=???$ is the Lamb-Dicke parameter for
optical pumping.
The geometry of the relavant beams (???and their polarizations) is showed in figure \ref{f-setup}B.
The optical tweezer has one weakly confined axial direction (axis $1$) and
two more strongly confined radial directions (axis $2$ and $3$).
Multiple pairs of beams are used to drive the Raman transition during cooling in order to
isolate and maximize the coupling to different trap axis.

Although Raman sideband cooling to vibrational ground state has already been successfully
implemented on neutral atoms in other experiments using heavier species like Rubidium [????]
and Cesium [????], such experiments generally use good polarization gradient cooling
and a tight confinement to achieve a small Lamb-Dicke parameter and effective Lamb-Dicke parameter.
However, this regime is harder to achieve with Sodium atoms, creating additional challenges for
us to realize ground state cooling. On the other hand, since these conditions can also be difficult
to meet for other interesting systemd like directly laser cooled molecules, the techniques we
use to overcome these challenges could be useful for a wider variety of systems.

\begin{figure}
  \includegraphics[width=9cm]{../../calculations/sideband_strength/imgs/coupling_0.46_0-6.png}
  \includegraphics[width=7cm]{imgs/heating_op.png}
  \caption{(A) Matrix elements as a function of $n$ and $\Delta n$ (showing complex dependency and zeros) (B) State changing probability. Large and increases as $n$ gets larger
    \label{f-ld}}
\end{figure}

As mentioned previously, in order to perform Raman sideband cooling efficiently and
minimize the heating during the optical pumping, we need to have a low initial temperature and
a small Lamb-Dicke parameter. However, since the Lamb-Dicke parameter $\eta$ is inversely
proportional to $\sqrt{m}$ (???) where $m$ is the mass of the atom, it is larger for Sodium
given the same trap depth (???). With $45mW$ of power at the focus of the trap,
we measured a trapping frequency of $\{\omega_1,\omega_2,\omega_3\}/2\pi = \{67, 420, 580\}\ \text{kHz}$
which corresponds to optical pumping Lamb-Dicke parameters of
$\{\eta^{OP}_1,\eta^{OP}_2,\eta^{OP}_3\} = \{???, ???, ???\}$.
Moreover, due to the unresolved(?) hyperfine structure in the $3^2P_{3/2}$ manifold,
the sub-doppler cooling in Sodium is also less efficient and we start the
Raman sideband cooling with a initial temperature of $40\mu K$. Combined with the high Lamb-Dicke
parameters, this gives us a initial effective optical pumping Lamb-Dicke parameters of
$\{\eta^{OP}_{1eff},\eta^{OP}_{2eff},\eta^{OP}_{3eff}\} = \{???, ???, ???\}$
As a result, there is a very high probability of heating during the optical pumping step
(figure \ref{f-ld}B) causing a $30\%$ average heating probability during optical pumping
in the weakly confined axial direction. Fortunately, the high Lamb-Dicke parameters also
provide us tools to overcome this issue. From our geometry, the Lamb-Dicke parameters for
Raman transitions are $\{\eta^R_{1},\eta^R_{2},\eta^R_{3}\} = \{???, ???, ???\}$. As shown in
figure \ref{f-ld}A, the high Raman Lamb-Dicke parameters causes a strong coupling to higher orders
of cooling sidebands, especially for high vibrational states.
This enables us to cool atoms in high vibrational states by driving on high order Raman sidebands,
removing more vibrational energy in a single cooling pulse and offsetting the effect of
strong heating. Since the heating probability is higher for high vibrational states,
cooling on these high order sidebands can greately suppress the high heating during
optical pumping during the initial cooling. Since the coupling strength of different orders
do not reach their minimums at the same time, using multiple orders of vibrational sidebands
for cooling also avoids accumulation of population near the coupling minimum of a particular
order, which improves the overall efficiency of the cooling process.

In additional to improving the cooling efficiency, the large Raman Lamb-Dicke parameters $\eta^R$
also creates difficulties for measuring the temperature. Traditional sideband thermometry uses
the ratio of the cooling and heating sidebands to measure $\bar n / (\bar n + 1)$, which relies
on the coupling strength to be proportional to $\sqrt{n}$. However, since we are not in the
Lamb-Dicke regime, the coupling strength diviates from this simple scaling rule quickly as already
shown in figure \ref{f-ld}A, causing the normal sideband thermometry to break down.
In order to solve this problem, when the atom temperature is still high,
we measure the heights of multiple sidebands to make sure no population is hidden at the
minimum of one sideband order. And when the atom is cooled down, we estimate the temperature
and ground state population under the assumption that only a few states are populated.
This result is then varified using independent measure of Rabi flopping on the carrier and heating
sidebands since they provide more information about the distribution of coupling strength.

Another issue caused by the high initial temperature is trap anharmonicity.
Although the potential near the center of a focused Gaussian beam can be approximated
by a harmonic trap, when a large number of levels are populated due to high initial temperature
the anharmonicity of the trap can lead to broadening of the sideband and the decrease of
the sideband signal. In our setup, this effect limits the Rabi frequency of the Raman pulse
on the radial sidebands to be no lower than tens of kilo-Hertz in order to drive atoms in
different vibrational states equally.

Finally, another important feature of our setup is the strong AC starck shift in the excited state.
The deep trap that is needed to trap and image single Sodium atom creates a light shift on the

the deep trap needed to ... also creates a large differential ..., which makes it impossible to load .... The same effect also happens on OP causing not only ... but also .... due to the breakdown of excited state HF structure.... We use switching....

% * Loose confinement
% ** More heating
% ** Complex sideband spectrum
% ** Matrix element minimum
% * Trap depth limit due to switching
% * High occupation
% ** Need more cooling
% ** Trap anharmonicity

% * Loose confinement
% ** Use higher orders to cool

Taking all the features of our system into account, we use a Monte-Carlo similation to verify
the validity of our method and explore the large parameter space for the cooling sequence.
In the simulation, we can observe the high heating rate due to the high Lamb-Dicke parameters
and confirm that high order of Raman sideband can be used to suppress this effect and reduces
the vibrational energy of the atom faster.
We also use the simulation to find a robust cooling strategy. As shown in figure (?),
we found that instead of cooling on only one sideband order at a time, it is generally more
efficient to alternate the cooling pulse between two neighboring orders. A cooling sequence
like this minimizes the accumulation of atom in a state not addressed by a particular sideband order.\\

\begin{figure*}
  \includegraphics[width=8cm]{imgs/spectrum_r.png}
  \includegraphics[width=8cm]{../../experiments/rabi_flop/imgs/fit_20170409_r2_0_ba.png}
  \caption{(A) (B) \label{f-radial}}
\end{figure*}

For the more tightly confined radial directions, we starts the cooling with $\{\bar n_2, \bar n_3\}=????, ????$. The sideband spectra of the initial distribution is shown in figure \ref{f-radial}A where we can clearly see the first order heating, first order cooling and second order cooling sidebands. After applying about 1000 cooling pulses cooling in all three dimentions starting with cooling on the radial second order, the Raman spectrum with the same parameter is shown in figure \ref{f-radial}B, where both cooling orders are suppressed. We can estimate the ground state probability in each direction based on the height of the first order cooling and heating sideband and the absence of the second order cooling sideband to be $......$. Due to the complexity of the sideband structure, we also use the Rabi flopping on the carrier and heating sideband as a independent measurement of the ground state population and gets $......$, showing good agreement between the two methods. (????? figure \ref{f-radial}C and \ref{f-radial}D)\\

\begin{figure*}
  \includegraphics[width=8cm]{imgs/spectrum_a1.png}
  \includegraphics[width=8cm]{../../experiments/rabi_flop/imgs/fit_20170409_a1_0_ba.png}
  \caption{(A) (B) \label{f-axial}}
\end{figure*}

As mentioned previously, the axial direction is much less confined and therefore has a much higher
effective Lamb-Dicke parameter of $\sqrt{n}\eta_{Raman1}\approx2.4$. The Raman spectrum before applying Raman sideban cooling is shown in figure \ref{f-axial}A where we can see clearly resolved Raman cooling sidebands up to the 8th order suggesting that we have many vibrational states populated in this direction.\\

\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\
\ \\

% Missing:
% ?? Switching helps OP fidelity
% ?? off-resonant scattering
% ?? D1 OP
% ?? Switching improves

\bibliography{paper}
\end{document}

% Generalization to other systems (properties that are shared with "more real" systems)
% Inharmonicity -> gaussian beam. -> importance to use short/strong pulses.