### 2017-07-31 presentations/2017-08-02_shutdown/checklist.tex
[presentations/2017-08-02_shutdown] WIP shutdown list

\documentclass[10pt,fleqn,twocolumn]{article}
% \usepackage[journal=rsc]{chemstyle}
% \usepackage{mhchem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{esint}
\usepackage{bbm}
\usepackage{amscd}
\usepackage{picinpar}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usepackage{indentfirst}
\usepackage{wrapfig}
\usepackage{units}
\usepackage{textcomp}
\usepackage[utf8x]{inputenc}
% \usepackage{feyn}
\usepackage{feynmp}
\usepackage{xkeyval}
\usepackage{xargs}
\usepackage{verbatim}
\usepackage{pgfplots}
\usepackage{hyperref}
\usetikzlibrary{
  arrows,
  calc,
  decorations.pathmorphing,
  decorations.pathreplacing,
  decorations.markings,
  fadings,
  positioning,
  shapes
}

\DeclareGraphicsRule{*}{mps}{*}{}
\newcommand{\ud}{\mathrm{d}}
\newcommand{\ue}{\mathrm{e}}
\newcommand{\ui}{\mathrm{i}}
\newcommand{\res}{\mathrm{Res}}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\dsum}{\displaystyle\sum}
\newcommand{\dprod}{\displaystyle\prod}
\newcommand{\dlim}{\displaystyle\lim}
\newcommand{\dint}{\displaystyle\int}
\newcommand{\fsno}[1]{{\!\not\!{#1}}}
\newcommand{\eqar}[1]
{
  \begin{align*}
    #1
  \end{align*}
}
\newcommand{\texp}[2]{\ensuremath{{#1}\times10^{#2}}}
\newcommand{\dexp}[2]{\ensuremath{{#1}\cdot10^{#2}}}
\newcommand{\eval}[2]{{\left.{#1}\right|_{#2}}}
\newcommand{\paren}[1]{{\left({#1}\right)}}
\newcommand{\lparen}[1]{{\left({#1}\right.}}
\newcommand{\rparen}[1]{{\left.{#1}\right)}}
\newcommand{\abs}[1]{{\left|{#1}\right|}}
\newcommand{\sqr}[1]{{\left[{#1}\right]}}
\newcommand{\crly}[1]{{\left\{{#1}\right\}}}
\newcommand{\angl}[1]{{\left\langle{#1}\right\rangle}}
\newcommand{\tpdiff}[4][{}]{{\paren{\frac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}_{#4}}}
\newcommand{\tpsdiff}[4][{}]{{\paren{\frac{\partial^{#1}}{\partial {#3}{}^{#1}}{#2}}_{#4}}}
\newcommand{\pdiff}[3][{}]{{\frac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}}
\newcommand{\diff}[3][{}]{{\frac{\ud^{#1} {#2}}{\ud {#3}{}^{#1}}}}
\newcommand{\psdiff}[3][{}]{{\frac{\partial^{#1}}{\partial {#3}{}^{#1}} {#2}}}
\newcommand{\sdiff}[3][{}]{{\frac{\ud^{#1}}{\ud {#3}{}^{#1}} {#2}}}
\newcommand{\tpddiff}[4][{}]{{\left(\dfrac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}\right)_{#4}}}
\newcommand{\tpsddiff}[4][{}]{{\paren{\dfrac{\partial^{#1}}{\partial {#3}{}^{#1}}{#2}}_{#4}}}
\newcommand{\pddiff}[3][{}]{{\dfrac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}}
\newcommand{\ddiff}[3][{}]{{\dfrac{\ud^{#1} {#2}}{\ud {#3}{}^{#1}}}}
\newcommand{\psddiff}[3][{}]{{\frac{\partial^{#1}}{\partial{}^{#1} {#3}} {#2}}}
\newcommand{\sddiff}[3][{}]{{\frac{\ud^{#1}}{\ud {#3}{}^{#1}} {#2}}}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{fontenc}
% \usepackage{tipa}
\usepackage{ulem}
\usepackage{color}
\usepackage{cancel}
\newcommand{\hcancel}[2][black]{\setbox0=\hbox{#2}%
  \rlap{\raisebox{.45\ht0}{\textcolor{#1}{\rule{\wd0}{1pt}}}}#2}
\pagestyle{fancy}
\setlength{\headheight}{67pt}
\fancyhead{}
\fancyfoot{}
\fancyfoot[C]{\thepage}
\fancyhead[R]{}
\renewcommand{\footruleskip}{0pt}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

\newcommand\pgfmathsinandcos[3]{%
  \pgfmathsetmacro#1{sin(#3)}%
  \pgfmathsetmacro#2{cos(#3)}%
}
\newcommand\LongitudePlane[3][current plane]{%
  \pgfmathsinandcos\sinEl\cosEl{#2} % elevation
  \pgfmathsinandcos\sint\cost{#3} % azimuth
  \tikzset{#1/.estyle={cm={\cost,\sint*\sinEl,0,\cosEl,(0,0)}}}
}
\newcommand\LatitudePlane[3][current plane]{%
  \pgfmathsinandcos\sinEl\cosEl{#2} % elevation
  \pgfmathsinandcos\sint\cost{#3} % latitude
  \pgfmathsetmacro\yshift{\cosEl*\sint}
  \tikzset{#1/.estyle={cm={\cost,0,0,\cost*\sinEl,(0,\yshift)}}} %
}
\newcommand\DrawLongitudeCircle[2][1]{
  \LongitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{atan(sin(#2)*cos(\angEl)/sin(\angEl))} %
  \draw[current plane] (\angVis:1) arc (\angVis:\angVis+180:1);
  \draw[current plane,dashed] (\angVis-180:1) arc (\angVis-180:\angVis:1);
}
\newcommand\DrawLatitudeCircleArrow[2][1]{
  \LatitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  \pgfmathsetmacro\sinVis{sin(#2)/cos(#2)*sin(\angEl)/cos(\angEl)}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{asin(min(1,max(\sinVis,-1)))}
  \draw[current plane,decoration={markings, mark=at position 0.6 with {\arrow{<}}},postaction={decorate},line width=.6mm] (\angVis:1) arc (\angVis:-\angVis-180:1);
  \draw[current plane,dashed,line width=.6mm] (180-\angVis:1) arc (180-\angVis:\angVis:1);
}
\newcommand\DrawLatitudeCircle[2][1]{
  \LatitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  \pgfmathsetmacro\sinVis{sin(#2)/cos(#2)*sin(\angEl)/cos(\angEl)}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{asin(min(1,max(\sinVis,-1)))}
  \draw[current plane] (\angVis:1) arc (\angVis:-\angVis-180:1);
  \draw[current plane,dashed] (180-\angVis:1) arc (180-\angVis:\angVis:1);
}
\newcommand\coil[1]{
  {\rh * cos(\t * pi r)}, {\apart * (2 * #1 + \t) + \rv * sin(\t * pi r)}
}
\makeatletter
\define@key{DrawFromCenter}{style}[{->}]{
  \tikzset{DrawFromCenterPlane/.style={#1}}
}
\define@key{DrawFromCenter}{r}[1]{
  \def\@R{#1}
}
\define@key{DrawFromCenter}{center}[(0, 0)]{
  \def\@Center{#1}
}
\define@key{DrawFromCenter}{theta}[0]{
  \def\@Theta{#1}
}
\define@key{DrawFromCenter}{phi}[0]{
  \def\@Phi{#1}
}
\presetkeys{DrawFromCenter}{style, r, center, theta, phi}{}
\newcommand*\DrawFromCenter[1][]{
  \setkeys{DrawFromCenter}{#1}{
    \pgfmathsinandcos\sint\cost{\@Theta}
    \pgfmathsinandcos\sinp\cosp{\@Phi}
    \pgfmathsinandcos\sinA\cosA{\angEl}
    \pgfmathsetmacro\DX{\@R*\cost*\cosp}
    \pgfmathsetmacro\DY{\@R*(\cost*\sinp*\sinA+\sint*\cosA)}
    \draw[DrawFromCenterPlane] \@Center -- ++(\DX, \DY);
  }
}
\newcommand*\DrawFromCenterText[2][]{
  \setkeys{DrawFromCenter}{#1}{
    \pgfmathsinandcos\sint\cost{\@Theta}
    \pgfmathsinandcos\sinp\cosp{\@Phi}
    \pgfmathsinandcos\sinA\cosA{\angEl}
    \pgfmathsetmacro\DX{\@R*\cost*\cosp}
    \pgfmathsetmacro\DY{\@R*(\cost*\sinp*\sinA+\sint*\cosA)}
    \draw[DrawFromCenterPlane] \@Center -- ++(\DX, \DY) node {#2};
  }
}
\makeatother
\tikzstyle{snakearrow} = [decorate, decoration={pre length=0.2cm,
  post length=0.2cm, snake, amplitude=.4mm,
  segment length=2mm},thick, ->]
%% document-wide tikz options and styles
\tikzset{%
  >=latex, % option for nice arrows
  inner sep=0pt,%
  outer sep=2pt,%
  mark coordinate/.style={inner sep=0pt,outer sep=0pt,minimum size=3pt,
    fill=black,circle}%
}
\addtolength{\hoffset}{-1.3cm}
\addtolength{\voffset}{-2cm}
\addtolength{\textwidth}{3cm}
\addtolength{\textheight}{2.5cm}
\renewcommand{\footskip}{10pt}
\setlength{\headwidth}{\textwidth}
\setlength{\headsep}{20pt}
\setlength{\marginparwidth}{0pt}

\makeatletter
\newcommand\ColorBox[2]{
  \begin{tikzpicture}
    \fill[#1] (0, 0) rectangle (\f@size/25, \f@size/25);
    \path (\f@size/50, \f@size/50) node[color=black!15] {\small #2};
  \end{tikzpicture}
}
\newcommand\DevBoxes{{\Large\ColorBox{color=red!50}{Off}\!\!\ColorBox{color=green!50}{On}}}
\newcommand\OffBox{{\normalsize\ColorBox{color=orange!50}{}}}
\newcommand\OnBox{{\normalsize\ColorBox{color=cyan!50}{}}}
\makeatother

\newcommand\DevHeader[1]{{\Large\DevBoxes\textbf{#1}\\}}
\def\arraystretch{2}

\parindent=0pt
\title{Checklist}
\date{August 2nd, 2017}
\begin{document}

\begin{itemize}
\item\DevHeader{1.0 Camera}
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Switch off
  \item\OffBox Switch off chiller
  \item\OffBox Unplug camera power (zip tie to camera)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in camera power
  \item\OnBox Switch on chiller
  \item\OnBox Switch on
  \end{itemize}
\item\DevHeader{1.5 Camera}
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Switch off
  \item\OffBox Switch off chiller
  \item\OffBox Unplug camera power (zip tie to camera)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in camera power
  \item\OnBox Switch on chiller
  \item\OnBox Switch on
  \end{itemize}
\item\DevHeader{Tweezer 1 computer}
  \textbf{Check following devices}
  \begin{itemize}
  \item\OnBox Thorlabs cameras
  \item\OnBox Na Raman amplifier
  \end{itemize}
\item\DevHeader{Tweezer 2 computer}
  \textbf{Check following devices}
  \begin{itemize}
  \item\OnBox Andor camera
  \end{itemize}
\item\DevHeader{Tweezer 3 computer}
  \textbf{Check following devices}
  \begin{itemize}
  \item\OnBox Software radio
  \end{itemize}
\item\DevHeader{Tweezer 4 computer}
  \textbf{Check following devices}
  \begin{itemize}
  \item\OnBox Andor camera
  \item\OnBox Thorlabs camera
  \end{itemize}
\item\DevHeader{1.0 Cs MOT PLL}
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug all signals (zip tie together)
  \item\OffBox Unplug 15V and 5V powers (zip tie together)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plugin 15V and 5V powers
  \item\OnBox Plugin all signals
  \item\OnBox Apply settings according to note on Generic tab, Equipment Settings.
  \end{itemize}
\item\DevHeader{1.0 Cs Raman PLL}
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug all signals (zip tie together)
  \item\OffBox Unplug 15V and 5V powers (zip tie together)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plugin 15V and 5V powers
  \item\OnBox Plugin all signals
  \item\OnBox Apply settings according to note on Generic tab, Equipment Settings.
  \end{itemize}
\item\DevHeader{1.5 Cs MOT PLL}
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug all signals (zip tie together)
  \item\OffBox Unplug 15V and 5V powers (zip tie together)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plugin 15V and 5V powers
  \item\OnBox Plugin all signals
  \item\OnBox Apply settings ???
  \end{itemize}
\item\DevHeader{1.5 Feshbach coils}
  \textbf{Before power shutdown}
  \begin{itemize}
  \item\OffBox Wrap exposed part in napkins to prevent condensations
  \end{itemize}
\item\DevHeader{1.5 IGBT}
  \textbf{Before power shutdown}
  \begin{itemize}
  \item\OffBox Wrap exposed part in napkins to prevent condensations
  \end{itemize}
\item\DevHeader{1.0 computer control box}
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox (Do following three steps quickly)
  \item\OffBox Turn off the box by flipping the switch on the front side of the box
  \item\OffBox Unplug the high voltage (48V) power supply (circular plug)
    on the front side of the box. Then unplug the power supply itself fromm the outlet.
  \item\OffBox Unplug the 12V board power supply.
    (L.T.E. one with a rectangular connector in the front)
  \item\OffBox Unplug USB power on the front side near the fan.
  \item\OffBox Zip tie the three unpluged powers to the computer control box.
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Connect to 3.5G Windfrek clock generator from Tweezer1.
    Set frequency to 3.5G and amplitude to max.
  \item\OnBox (Remove power connection zip ties and) Plug in USB power supply.
  \item\OnBox (Do following three steps quickly)
  \item\OnBox Plug in the 12V board power supply.
  \item\OnBox Plug in high voltage power supply.
  \item\OnBox Turn on the power switch of the box. The fans should start spinning.
  \end{itemize}
\item\DevHeader{1.5 computer control box}
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox (Do following three steps quickly)
  \item\OffBox Turn off the box by flipping the switch on the front side of the box
  \item\OffBox Unplug the high voltage (48V) power supply (circular plug)
    on the left side of the box. Then unplug the power supply itself fromm the outlet.
  \item\OffBox Unplug the 12V board power supply.
    (L.T.E. one with a rectangular connector in the front)
  \item\OffBox Unplug USB power on the front side near the fan.
  \item\OffBox Zip tie the three unpluged powers to the computer control box.
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Connect to 3.5G Windfrek clock generator from Tweezer1.
    Set frequency to 3.5G and amplitude to max.
  \item\OnBox (Remove power connection zip ties and) Plug in USB power supply.
  \item\OnBox (Do following three steps quickly)
  \item\OnBox Plug in the 12V board power supply.
  \item\OnBox Plug in high voltage power supply. The fans should start spinning.
  \item\OnBox Turn on the power switch of the box.
  \end{itemize}
\item\DevHeader{Cs 1.0 MOT home built temperature servo}
  \textbf{1st level above Cs tweezer/MOT lasers}\\
  Switch off/on the front switch
\item\DevHeader{Benchtop power supply}
  \textbf{1st level above Cs tweezer/MOT lasers}\\
  Record voltage/current values:\\
  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
\item\DevHeader{1st version Till's current controller for Cs MOT}
  \textbf{1st level above Cs tweezer/MOT lasers}\\
  Record current value:\\
  \begin{tabular}{|c|c|}
    \hline
    Current&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power (zip tie to controller)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Turn on
  \end{itemize}
\item\DevHeader{Homebuilt $\pm15,5$V power supply}
  \textbf{2nd level above Cs tweezer/MOT lasers}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
\end{itemize}

\end{document}