### 2018-06-20 designs/pre-amp/notes.tex
[designs/pre-amp] more notes

\documentclass[10pt,fleqn]{article}

\usepackage{hyperref}

\begin{document}

\section{Value selection for active low pass filter}
\begin{tabular}{|c|c|c|c|c|c|c|}
  \hline
  Gain&$f_{cutoff}\ (\mathrm{kHz})$&$R_5\ (\mathrm{k\Omega})$&$R_6\ (\mathrm{k\Omega})$&$R_7\ (\mathrm{k\Omega})$&$C_{12}\ (\mathrm{pF})$&$C_{13}\ (\mathrm{pF})$\\\hline
  1&10&1.54&1.54&11.1&10000&910\\\hline
  1&20&4.99&5.0&4.64&2200&750\\\hline
  1&30&1.01&1.01&1.1&6800&2200\\\hline
  0.5&10&5.0&2.5&17.4&4700&750\\\hline
  0.5&20&2.0&1.01&10.7&5600&620\\\hline
  0.5&30&3.88&1.96&6.34&2200&620\\\hline
  0.25&10&2.21&0.562&18.7&15000&820\\\hline
  0.25&20&2.7&0.681&10.7&6800&750\\\hline
  0.25&30&1.82&0.464&6.57&6800&820\\\hline
\end{tabular}
The active filter uses a low-pass multiple-feedback (MFB) architecture
and the table gives the values to create a second order Bessel filter.
All the resistors and capacitors are currently in stock at Digikey.
See \href{http://www.ti.com/lit/an/sloa049b/sloa049b.pdf}
{Active Low-Pass Filter Design}\footnote{http://www.ti.com/lit/an/sloa049b/sloa049b.pdf}
for the formula used and comparison between different filter type and circuit topology.

\section{Other tweaks}
\begin{enumerate}
\item $R_8$ and $C_{14}$ can be used for additional filter.
\item If a higher cutoff frequency is needed, reduce or remove $C_2$ to increase the
  input cutoff frequency.
\end{enumerate}

\end{document}