### 2019-04-26 presentations/2019-04-27_shutdown/checklist.tex
[presentation/2019-04-27_shutdown] import equipment list

\documentclass[10pt,fleqn,twocolumn]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{esint}
\usepackage{bbm}
\usepackage{amscd}
\usepackage{picinpar}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usepackage{indentfirst}
\usepackage{wrapfig}
\usepackage{units}
\usepackage{textcomp}
\usepackage[utf8x]{inputenc}
% \usepackage{feyn}
\usepackage{feynmp}
\usepackage{xkeyval}
\usepackage{xargs}
\usepackage{verbatim}
\usepackage{pgfplots}
\usepackage{hyperref}
\usepackage{longtable}
% \makeatletter
% \let\oldlt\longtable
% \let\endoldlt\endlongtable
% \def\longtable{\@ifnextchar[\longtable@i \longtable@ii}
% \def\longtable@i[#1]{\begin{figure}[t]
% \onecolumn
% \begin{minipage}{0.5\textwidth}
% \oldlt[#1]
% }
% \def\longtable@ii{\begin{figure}[t]
% \onecolumn
% \begin{minipage}{0.5\textwidth}
% \oldlt
% }
% \def\endlongtable{\endoldlt
% \end{minipage}
% \twocolumn
% \end{figure}}
% \makeatother

\usetikzlibrary{
  arrows,
  calc,
  decorations.pathmorphing,
  decorations.pathreplacing,
  decorations.markings,
  fadings,
  positioning,
  shapes
}

\DeclareGraphicsRule{*}{mps}{*}{}
\newcommand{\ud}{\mathrm{d}}
\newcommand{\ue}{\mathrm{e}}
\newcommand{\ui}{\mathrm{i}}
\newcommand{\res}{\mathrm{Res}}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\dsum}{\displaystyle\sum}
\newcommand{\dprod}{\displaystyle\prod}
\newcommand{\dlim}{\displaystyle\lim}
\newcommand{\dint}{\displaystyle\int}
\newcommand{\fsno}[1]{{\!\not\!{#1}}}
\newcommand{\eqar}[1]
{
  \begin{align*}
    #1
  \end{align*}
}
\newcommand{\texp}[2]{\ensuremath{{#1}\times10^{#2}}}
\newcommand{\dexp}[2]{\ensuremath{{#1}\cdot10^{#2}}}
\newcommand{\eval}[2]{{\left.{#1}\right|_{#2}}}
\newcommand{\paren}[1]{{\left({#1}\right)}}
\newcommand{\lparen}[1]{{\left({#1}\right.}}
\newcommand{\rparen}[1]{{\left.{#1}\right)}}
\newcommand{\abs}[1]{{\left|{#1}\right|}}
\newcommand{\sqr}[1]{{\left[{#1}\right]}}
\newcommand{\crly}[1]{{\left\{{#1}\right\}}}
\newcommand{\angl}[1]{{\left\langle{#1}\right\rangle}}
\newcommand{\tpdiff}[4][{}]{{\paren{\frac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}_{#4}}}
\newcommand{\tpsdiff}[4][{}]{{\paren{\frac{\partial^{#1}}{\partial {#3}{}^{#1}}{#2}}_{#4}}}
\newcommand{\pdiff}[3][{}]{{\frac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}}
\newcommand{\diff}[3][{}]{{\frac{\ud^{#1} {#2}}{\ud {#3}{}^{#1}}}}
\newcommand{\psdiff}[3][{}]{{\frac{\partial^{#1}}{\partial {#3}{}^{#1}} {#2}}}
\newcommand{\sdiff}[3][{}]{{\frac{\ud^{#1}}{\ud {#3}{}^{#1}} {#2}}}
\newcommand{\tpddiff}[4][{}]{{\left(\dfrac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}\right)_{#4}}}
\newcommand{\tpsddiff}[4][{}]{{\paren{\dfrac{\partial^{#1}}{\partial {#3}{}^{#1}}{#2}}_{#4}}}
\newcommand{\pddiff}[3][{}]{{\dfrac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}}
\newcommand{\ddiff}[3][{}]{{\dfrac{\ud^{#1} {#2}}{\ud {#3}{}^{#1}}}}
\newcommand{\psddiff}[3][{}]{{\frac{\partial^{#1}}{\partial{}^{#1} {#3}} {#2}}}
\newcommand{\sddiff}[3][{}]{{\frac{\ud^{#1}}{\ud {#3}{}^{#1}} {#2}}}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{fontenc}
\usepackage{titlesec}
% \usepackage{tipa}
\usepackage{ulem}
\usepackage{color}
\usepackage{cancel}

\newcommand{\hcancel}[2][black]{\setbox0=\hbox{#2}%
  \rlap{\raisebox{.45\ht0}{\textcolor{#1}{\rule{\wd0}{1pt}}}}#2}
\pagestyle{fancy}
% \setlength{\headheight}{60pt}
\fancyhead{}
\fancyfoot{}
\fancyfoot[C]{\thepage}
\fancyhead[R]{}
\renewcommand{\footruleskip}{0pt}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}
\titlespacing{\subsection}{0pt}{\parskip}{-\parskip}

\newcommand\pgfmathsinandcos[3]{%
  \pgfmathsetmacro#1{sin(#3)}%
  \pgfmathsetmacro#2{cos(#3)}%
}
\newcommand\LongitudePlane[3][current plane]{%
  \pgfmathsinandcos\sinEl\cosEl{#2} % elevation
  \pgfmathsinandcos\sint\cost{#3} % azimuth
  \tikzset{#1/.estyle={cm={\cost,\sint*\sinEl,0,\cosEl,(0,0)}}}
}
\newcommand\LatitudePlane[3][current plane]{%
  \pgfmathsinandcos\sinEl\cosEl{#2} % elevation
  \pgfmathsinandcos\sint\cost{#3} % latitude
  \pgfmathsetmacro\yshift{\cosEl*\sint}
  \tikzset{#1/.estyle={cm={\cost,0,0,\cost*\sinEl,(0,\yshift)}}} %
}
\newcommand\DrawLongitudeCircle[2][1]{
  \LongitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{atan(sin(#2)*cos(\angEl)/sin(\angEl))} %
  \draw[current plane] (\angVis:1) arc (\angVis:\angVis+180:1);
  \draw[current plane,dashed] (\angVis-180:1) arc (\angVis-180:\angVis:1);
}
\newcommand\DrawLatitudeCircleArrow[2][1]{
  \LatitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  \pgfmathsetmacro\sinVis{sin(#2)/cos(#2)*sin(\angEl)/cos(\angEl)}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{asin(min(1,max(\sinVis,-1)))}
  \draw[current plane,decoration={markings, mark=at position 0.6 with {\arrow{<}}},postaction={decorate},line width=.6mm] (\angVis:1) arc (\angVis:-\angVis-180:1);
  \draw[current plane,dashed,line width=.6mm] (180-\angVis:1) arc (180-\angVis:\angVis:1);
}
\newcommand\DrawLatitudeCircle[2][1]{
  \LatitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  \pgfmathsetmacro\sinVis{sin(#2)/cos(#2)*sin(\angEl)/cos(\angEl)}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{asin(min(1,max(\sinVis,-1)))}
  \draw[current plane] (\angVis:1) arc (\angVis:-\angVis-180:1);
  \draw[current plane,dashed] (180-\angVis:1) arc (180-\angVis:\angVis:1);
}
\newcommand\coil[1]{
  {\rh * cos(\t * pi r)}, {\apart * (2 * #1 + \t) + \rv * sin(\t * pi r)}
}
\makeatletter
\define@key{DrawFromCenter}{style}[{->}]{
  \tikzset{DrawFromCenterPlane/.style={#1}}
}
\define@key{DrawFromCenter}{r}[1]{
  \def\@R{#1}
}
\define@key{DrawFromCenter}{center}[(0, 0)]{
  \def\@Center{#1}
}
\define@key{DrawFromCenter}{theta}[0]{
  \def\@Theta{#1}
}
\define@key{DrawFromCenter}{phi}[0]{
  \def\@Phi{#1}
}
\presetkeys{DrawFromCenter}{style, r, center, theta, phi}{}
\newcommand*\DrawFromCenter[1][]{
  \setkeys{DrawFromCenter}{#1}{
    \pgfmathsinandcos\sint\cost{\@Theta}
    \pgfmathsinandcos\sinp\cosp{\@Phi}
    \pgfmathsinandcos\sinA\cosA{\angEl}
    \pgfmathsetmacro\DX{\@R*\cost*\cosp}
    \pgfmathsetmacro\DY{\@R*(\cost*\sinp*\sinA+\sint*\cosA)}
    \draw[DrawFromCenterPlane] \@Center -- ++(\DX, \DY);
  }
}
\newcommand*\DrawFromCenterText[2][]{
  \setkeys{DrawFromCenter}{#1}{
    \pgfmathsinandcos\sint\cost{\@Theta}
    \pgfmathsinandcos\sinp\cosp{\@Phi}
    \pgfmathsinandcos\sinA\cosA{\angEl}
    \pgfmathsetmacro\DX{\@R*\cost*\cosp}
    \pgfmathsetmacro\DY{\@R*(\cost*\sinp*\sinA+\sint*\cosA)}
    \draw[DrawFromCenterPlane] \@Center -- ++(\DX, \DY) node {#2};
  }
}
\makeatother
\tikzstyle{snakearrow} = [decorate, decoration={pre length=0.2cm,
  post length=0.2cm, snake, amplitude=.4mm,
  segment length=2mm},thick, ->]
%% document-wide tikz options and styles
\tikzset{%
  >=latex, % option for nice arrows
  inner sep=0pt,%
  outer sep=2pt,%
  mark coordinate/.style={inner sep=0pt,outer sep=0pt,minimum size=3pt,
    fill=black,circle}%
}
\addtolength{\hoffset}{-1.3cm}
\addtolength{\voffset}{-2cm}
\addtolength{\textwidth}{3cm}
\addtolength{\textheight}{2.5cm}
\renewcommand{\footskip}{10pt}
\setlength{\headwidth}{\textwidth}
\setlength{\headsep}{20pt}
\setlength{\marginparwidth}{0pt}

\makeatletter
\newcommand\ColorBox[2]{
  \begin{tikzpicture}
    \fill[#1] (0, 0) rectangle (\f@size/25, \f@size/25);
    \draw[color=gray!80] (0, 0) rectangle (\f@size/25, \f@size/25);
    \path (\f@size/50, \f@size/50) node[color=black!15] {\small #2};
  \end{tikzpicture}
}
\newcommand\DevBoxes{\mbox{{\Large\ColorBox{color=red!35}{Off}\!\!\ColorBox{color=green!35}{On}}}}
\newcommand\OffBox{{\normalsize\ColorBox{color=orange!40}{}}}
\newcommand\OnBox{{\normalsize\ColorBox{color=cyan!40}{}}}
\makeatother

\newcommand\DevHeader[1]{{\Large\DevBoxes\textbf{#1}\\}}
\def\arraystretch{2}

\newcommand\DevDef[2]{
  \begin{minipage}[t]{\linewidth}
    \begin{itemize}
    \item\DevHeader{#1}
      #2
    \end{itemize}
    \vspace{0.8cm}
  \end{minipage}
}
\newcommand\Person[2]{
\clearpage
\section*{#1}
\fancyhead[R]{#1}
\setcounter{page}{1}
\setcounter{subsection}{0}
For shutting off, go through the checklist in the following order
\begin{itemize}
\item\OffBox Turn off equipments
\item\OffBox Turn off power supplies
\item\OffBox Turn off computers
\end{itemize}
Wait before everyone else finish their corresponding steps before starting the next one.\\
For powering up, go through the checklist in the following order
\begin{itemize}
\item\OnBox Turn on computers
\item\OnBox Turn on power supplies
\item\OnBox Turn on equipments
\item\OnBox Check computer connection to devices
\end{itemize}
Also wait before everyone else finish their corresponding steps before starting the next one.\\

\textbf{\Large #2}
\vfill
}
\renewcommand{\thesubsection}{\arabic{subsection}}

\parindent=0pt
\title{Checklist}
\date{August 2nd, 2017}
\begin{document}

\onecolumn
\Person{Yichao}{
  Equipments and power supplies are mostly located around the laser table on the library side.\\
  Check Lee's part after everything are off.
}
\begin{longtable}{|c|c|c|}
  \hline
  \textbf{Equipment}&\textbf{Location}&\textbf{Notes}\\\hline
  \multicolumn{3}{|c|}{\textbf{Laser table facing computer or the wall} west $\rightarrow$ south}\\\hline
  Benchtop power for Cs MOT beat lock&Lv1 above 1.0 Cs MOT&\\\hline
  PLL for Cs MOT&Below Cs MOT&Save setting\\\hline
  Home power (+-15,5)&Lv2 above Cs MOT&\\\hline
  Valan&1.5 amplifier stack Lv1 above 1.0 Cs MOT&Save setting\\\hline
  Till driver v2 for 1.5 Na Raman&Lv1 above 1.5 Na Raman&\\\hline
  Temp servo for 1.5 Na Raman seed&Lv1 above 1.5 Na Raman&\\\hline
  Temp servo for 1.5 Na Raman doubler&Lv1 above 1.5 Na Raman&\\\hline
  PLL for Cs Raman&Next to 1.0 Cs Raman&Save setting\\\hline
  New Focus driver for Cs Raman&Lv1 above 1.0 Cs Raman&Save setting\\\hline
  Benchtop power (x3)&Lv2 above 1.0 Cs MOT&\\\hline
  Home power (24V) x2&Lv2 above 1.0 Cs Raman&\\\hline
  Benchtop power&Lv1 next to New Focus driver&\\\hline
  Shutter driver&Lv1 above 1.0 Cs Raman&\\\hline
  New Focus lock box for 1.0 Cs Raman&Lv1 above 1.0 Cs Raman&\\\hline
  Thorlabs temp servo for 1.0 Na Raman doubler&Lv1 above 1.0 Na seed&\\\hline
  Timebase drivers for 1.0 Na Raman and Na D2&Lv1 above Thorlabs temp servo&Save setting\\\hline
  Benchtop power for 1.0 Cs Raman&Lv2 above New Focus lock box&\\\hline
  High voltage power&Lv2 above timebase drivers&\\\hline
  Home lock box for Na D2&Lv1 above 1.0 Na seed&\\\hline
  MPB Raman laser&Lv1 above 1.0 Na MOT&Save setting\\\hline
  Oscilloscope for Na D2 lock&Lv1 above MPB Raman laser&\\\hline
  Benchtop power for 1.0 Na Raman seed&Lv2 above oscilloscope&\\\hline
\end{longtable}
\twocolumn

\subsection{Equipments}
\DevDef{Thorlabs temperature controller for Na Raman doubler}{
  \textbf{1st level above Na Raman and D2 seed lasers}\\
  Record temperature settings\\

  \begin{tabular}{|c|c|}
    \hline
    Setpoint&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Check temperature setting
  \item\OnBox Turn on
  \end{itemize}
}
\DevDef{Timebase driver for Na Raman and Na D2 (x2)}{
  \textbf{1st level above Na Raman and D2 seed lasers}\\
  Record current and T1 settings\\

  \begin{tabular}{|c|c|}
    \hline
    Raman Current&\hphantom{\hspace{3.5cm}}\\\hline
    Raman T1&\\\hline
    MOT Current&\hphantom{\hspace{3.5cm}}\\\hline
    MOT T1&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Turn on
  \item\OnBox Set current/temperature.
  \item\OnBox Enable T1 survo
  \item\OnBox Turn on current
  \end{itemize}
}
\DevDef{Oscilloscope for Na D2 lock}{
  \textbf{1st level above Na Raman amplifier}
  \begin{itemize}
  \item\OffBox Unplug from wall
  \end{itemize}
}
\DevDef{MPB Raman amplifier}{
  \textbf{1st level above Na MOT beam path}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Power off
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into wall
  \item\OnBox Power on
  \end{itemize}
}
\DevDef{Homebuilt lockbox for Na D2}{
  \textbf{1st level above Na seed}\\
  \textbf{Turn on-off together with the adjustable high voltage power supply above}\\
  Record high voltage power supply output

  \begin{tabular}{|c|c|}
    \hline
    Voltage&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unlock
  \item\OffBox Turn down the voltage of the high voltage power supply to 0
  \item\OffBox Unplug high voltage power on lockbox (zip tie)
  \item\OffBox Unplug 15V from lockbox (zip tie)
  \item\OffBox Turn off high voltage power supply
  \item\OffBox Unplug high voltage power supply from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug high voltage power supply into the wall
  \item\OnBox Turn on high voltage power supply (make sure the output is 0)
  \item\OnBox Plug in 15V to lockbox
  \item\OnBox Plug in high voltage to lockbox
  \item\OnBox Turn up the voltage of the high voltage power supply
  \item\OnBox Check HV output
  \end{itemize}
}
\DevDef{SRS shutter driver for Na}{
  \textbf{1st level above Na D2 beam path, below Variac}\\
  Switch off/on with the switch on the back.
}
\DevDef{Variac for Na cell}{
  \textbf{Above Na D2 beam path}\\
  Record temperature and voltage settings\\

  \begin{tabular}{|c|c|}
    \hline
    Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Temperature&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn voltage to 0
  \item\OffBox Switch off
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into wall
  \item\OnBox Switch on
  \item\OnBox Turn the voltage up \textbf{SLOWLY} and monitor the temperature at the same time.
  \end{itemize}
}
\DevDef{Homebuilt lockbox for Na D1}{
  \textbf{1st level above Na D2 beam path}\\
  \textbf{Turn on-off together with the fixed high voltage power supply above}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unlock
  \item\OffBox Unplug high voltage power on lockbox (zip tie)
  \item\OffBox Unplug 15V from lockbox (zip tie)
  \item\OffBox Unplug high voltage power supply from walls
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug high voltage power supply into the wall
  \item\OnBox Plug in 15V to lockbox
  \item\OnBox Plug in high voltage to lockbox
  \item\OnBox Check HV output
  \end{itemize}
}
\DevDef{Thorlabs temperature controller for Na D1 doubler}{
  \textbf{1st level above Na D1}\\
  Record temperature settings\\

  \begin{tabular}{|c|c|}
    \hline
    Setpoint&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Check temperature setting
  \item\OnBox Turn on
  \end{itemize}
}
\DevDef{Thorlabs temperature controller for Na D1 seed}{
  \textbf{1st level above Na D1 seed}\\
  Record temperature settings\\

  \begin{tabular}{|c|c|}
    \hline
    Setpoint&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Check temperature setting
  \item\OnBox Turn on
  \end{itemize}
}
\DevDef{Thorlabs current controller for Na D1 seed}{
  \textbf{1st level above Na D1 seed}\\
  Record current settings\\

  \begin{tabular}{|c|c|}
    \hline
    Current&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Check current setting
  \item\OnBox Turn on
  \end{itemize}
}
\subsection{Power supplies}
\DevDef{Benchtop power supply}{
  \textbf{2nd level above Na Variac}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Homebuilt $24$V power supply}{
  \textbf{2nd level above Cs tweezer controllers}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Benchtop power supply}{
  \textbf{2nd level above Na D1 seed controllers}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Benchtop power supply}{
  \textbf{2nd level above Cs tweezer, right of 24V power supply}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Benchtop power supply}{
  \textbf{1st level above Cs tweezer}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{UPS}{
  \textbf{2nd level above laser table}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on
  \end{itemize}
}
\subsection{Computers}
\DevDef{Computer control box for 1.0}{
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox (Do following three steps quickly)
  \item\OffBox Turn off the box by flipping the switch on the front side of the box
  \item\OffBox Unplug the high voltage (48V) power supply (circular plug)
    on the front side of the box. Then unplug the power supply itself fromm the outlet.
  \item\OffBox Unplug the 12V board power supply.
    (L.T.E. one with a rectangular connector in the front)
  \item\OffBox Unplug USB power on the front side near the fan.
  \item\OffBox Zip tie the three unpluged powers to the computer control box.
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Connect to 3.5G Windfrek clock generator from Tweezer1.
    Set frequency to 3.5G and amplitude to max.
  \item\OnBox (Remove power connection zip ties and) Plug in USB power supply.
  \item\OnBox (Do following three steps quickly)
  \item\OnBox Plug in the 12V board power supply.
  \item\OnBox Plug in high voltage power supply.
  \item\OnBox Turn on the power switch of the box. The fans should start spinning.
  \end{itemize}
}

\onecolumn
\Person{Kenneth}{
  Equipments and power supplies are mostly located around the machine table facing the laser table
  or the computer area.\\
  Check Jessie's part after everything are off.
}
\begin{longtable}{|c|c|c|}
  \hline
  \textbf{Equipment}&\textbf{Location}&\textbf{Notes}\\\hline
  \multicolumn{3}{|c|}{\textbf{Laser table side facing machine table} north $\rightarrow$ east}\\\hline
  671 EOM stack&Floor&\\\hline
  1550 Amp&Floor&\\\hline
  Radio computer&Lv2&\\\hline
  Benchtop power&Lv2 next to radio computer&\\\hline
  High voltage power&Lv2 next to benchtop power&\\\hline
  Variac for Na cell&Lv1 above 1.0 Na MOT&\\\hline
  Shutter driver&Lv1 next to variac&\\\hline
  Home lock box for 1.0 Na D1&Lv1 next to shutter driver&\\\hline
  1038 Amp&Floor&\\\hline
  1038 Timebase laser&Table&\\\hline
  FPGA box for 1.0&Lv2&\\\hline
  Thorlabs temp servo for Na D1 doubler&Lv1 above Na D1&\\\hline
  Thorlabs temp servo for Na D1 seed&Lv1 above Na D1&\\\hline
  Thorlabs current for Na D1 seed&Lv1 above Na D1&\\\hline
  Benchtop power&Lv2 above Thorlabs controllers for Na D1 seed&\\\hline
  Home power (24)&Lv2 next to benchtop power&\\\hline
  Benchtop power x2&Lv2 next to Home power (24)&\\\hline
  Till driver v2 for 976&Lv1 above Na D1 doubler&\\\hline
  Thorlabs temp servo for Cs tweezer&Lv1 above Timebase 1038 beam&\\\hline
  Thorlabs current (2A) (off)&Lv1 above Thorlabs temp servo&\\\hline
  Till driver v2 for 671&Lv1 above Thorlabs current&\\\hline
  Home temp servo x2&Lv1 above Timebase 1038 control box&\\\hline
  Oscilloscope&Lv1&\\\hline
  \multicolumn{3}{|c|}{\textbf{Machine table side facing laser table} south}\\\hline
  Thorlabs current for 1.5 Cs MOT&Lv1 above 1.5 Cs&\\\hline
  Till driver v2 for 1.5 Cs RP&Lv1 above 1.5 Cs&\\\hline
  Home temp servo (x2) for 1.5 Cs MOT/RP&Lv1 above 1.5 Cs&\\\hline
  Oscilloscope for 1.5 Cs&Lv1 above 1.5 Cs&\\\hline
  Home lock box for 1.5 Cs&Lv1 above 1.5 Cs&\\\hline
  Novatec for 1.5 Cs lock&Lv1 above 1.5 Cs&Save setting\\\hline
  Benchtop power&Lv2 1.5 Cs MOT/RP current&\\\hline
  Home power (24) x2&Lv2 next to benchtop power&\\\hline
  Benchtop power&Lv2 on two home power (24)&\\\hline
  Home power (+-15,5)&Lv2 middle&\\\hline
  Benchtop power (x2) for uWave Amp&Lv2 next to home power (+-15,5)&\\\hline
  Benchtop power (x2)&Lv1 above 1.5 Na&\\\hline
  High voltage power&Lv1 next to 2 benchtop power&\\\hline
  Function generator for 1.0 MOT piezo&Lv1&\\\hline
  Temperature and pump controllers for cavity&Lv1 next to function generator&\\\hline
  Water chiller for uWave breadboard&Floor&\\\hline
  SAES pump&Lv1 above 1.0 chamber&\\\hline
  Benchtop power x2&Lv2 next to coil servo&\\\hline
\end{longtable}
\twocolumn
\subsection{Equipments}
\DevDef{Agilent function generator for Na 1.0 switching}{
  \textbf{1st level above TiSapph facing computers}\\
  Record parameters:\\

  \begin{tabular}{|c|c|}
    \hline
    Ch1 Hi&\hphantom{\hspace{3.5cm}}\\\hline
    Ch1 Lo&\\\hline
    Ch2 Hi&\\\hline
    Ch2 Lo&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into wall
  \item\OnBox Turn on
  \item\OnBox Set frequency, duty cycle, high/low voltages
  \item\OnBox Set switching phase back
  \end{itemize}
}
\DevDef{2nd version Till's current controller for Cs 1.5 (x2)}{
  \textbf{1st level above 1.5 Cs MOT}\\
  Record current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Top Current&\hphantom{\hspace{3.5cm}}\\\hline
    Bottom Current&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power (zip tie to controller)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Turn on
  \end{itemize}
}
\DevDef{Thorlabs current controller}{
  \textbf{1st level below Cs 1.5 MOT drivers}\\
  Unused\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power
  \end{itemize}
}
\DevDef{Homebuilt temperature servo for Cs 1.5 MOT (x2)}{
  \textbf{1st level above Cs 1.5 MOT}\\
  Switch off/on the front switch
}
\DevDef{Oscilloscope for Cs 1.5 MOT lock}{
  \textbf{1st level above Cs 1.5 MOT}
  \begin{itemize}
  \item\OffBox Unplug from wall
  \end{itemize}
}
\DevDef{Homebuilt lockbox for Cs 1.5 MOT}{
  \textbf{1st level above Cs 1.5 MOT}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unlock
  \item\OffBox Unplug 15V from lockbox (zip tie)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in 15V to lockbox
  \end{itemize}
}
\DevDef{Function generator for Cs MOT 1.5 lockbox}{
  \textbf{1st level under Cs MOT 1.5 lockbox}
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Record settings
  \item\OffBox Turn off
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into wall
  \item\OnBox Turn on
  \item\OnBox Restore settings
  \end{itemize}
}
\DevDef{HV amplifiers for 1.0 piezo mirrors}{
  \textbf{Above the space beween Na 1.5 beam path and NaCs 1.5 chamber}\\
  \textbf{Turn on-off together with the function generator below
    and the high voltage power supply on the left}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off the function generator
  \item\OffBox Unplug function generator output (zip tie and label)
  \item\OffBox Unplug HV power supply from the wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plugin HV power supply to the wall
  \item\OnBox Plugin function generator output
  \item\OnBox Turn on the function generator
  \end{itemize}
}
\DevDef{Pump}{
  \textbf{1st level above NaCs 1.0 chamber}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug from walls
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into wall
  \item\OnBox Turn on (should be enough)
  \end{itemize}
}
\DevDef{Andor Camera for 1.0}{
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Switch off
  \item\OffBox Switch off chiller
  \item\OffBox Unplug camera power (zip tie to camera)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in camera power
  \item\OnBox Switch on chiller
  \item\OnBox Switch on
  \end{itemize}
}
\subsection{Power supplies}
\DevDef{Benchtop power supply}{
  \textbf{2nd level above Cs 1.5 MOT controllers}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Homebuilt $24$V power supply}{
  \textbf{2nd level above Cs 1.5 MOT controllers}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Benchtop power supply (x2)}{
  \textbf{1st level above Na 1.5 bean path}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Homebuilt $\pm15,5$V power supply}{
  \textbf{2nd level above Na 1.5 beam path}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Benchtop power supply (x2)}{
  \textbf{2nd level above piezo mirror electronics}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Benchtop power supply}{
  \textbf{2nd level above NaCs 1.0 chamber}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\subsection{Computers}
\DevDef{Tweezer 2 computer}{
  \textbf{Check following devices}
  \begin{itemize}
  \item\OnBox Andor camera
  \end{itemize}
}
\DevDef{Tweezer 3 computer}{
  \textbf{Check following devices}
  \begin{itemize}
  \item\OnBox Software radio
  \end{itemize}
}

\onecolumn
\Person{Jessie}{
  Equipments and power supplies are mostly located around the machine table facing the 1.5 table.\\
  Check Frederic's part after everything are off.
}
\begin{longtable}{|c|c|c|}
  \hline
  \textbf{Equipment}&\textbf{Location}&\textbf{Notes}\\\hline
  \textbf{Machine table Facing the wall}&north&\\\hline
  Benchtop power&Lv2&\\\hline
  Home power (24)&Lv2 above 1.0 tweezer breadboard&\\\hline
  Home power (+-15,5)&Lv2 above home power (24)&\\\hline
  Benchtop power x9&Lv2 next to home power (24) and (+-15,5)&\\\hline
  Current servo for compensation coil&Lv1 below home power (24)&\\\hline
  Kenneth servo for 1.0 Cs tweezer&Lv1&\\\hline
  High voltage Amp for 630 lock&Lv1&\\\hline
  Vincent servo x2 for STIRAP lasers&Lv1&\\\hline
  Greiner servo for 1.0 Na tweezer&Lv1 above Vincent servos&\\\hline
  Rack for STRAP&Lv1 next to Vincent servos&\\\hline
  Till driver v2 x2 for STIRAP lasers&Lv1 above rack for 1.5&\\\hline
  Home temp servo x3 for STIRAP&Lv2 next to 9x benchtop power&\\\hline
  FPGA box for 1.5&Lv2&\\\hline
  Home power (+-15,5)&Lv2 next to FPGA box&\\\hline
  Benchtop power x2&Lv2 on home power (+-15,5)&\\\hline
  Benchtop power&Lv2 next to home power (+-15,5)&\\\hline
  Benchtop power&Floor rack Lv1&\\\hline
  Till driver v2 x2 for 1.5 Cs Raman&Floor rack Lv2&\\\hline
  Kenneth servo for 1.5 Cs Raman phase lock&Floor rack Lv2 above Till drivers&\\\hline
  Benchtop power&Floor rack Lv2 next to Till drivers&\\\hline
  Home temp servo x2 for 1.5 Cs Raman&Floor rack Lv2&\\\hline
  Piezo driver for 1.5 Cs Raman&Floor rack Lv2 above home temp servos&\\\hline
  Delay generator for 1.5 Na&Lv1 below 1.5 output panel&Save setting\\\hline
  Till driver v2 for 1.0 Innolume&Lv1 next to delay generator&\\\hline
  Kenneth servo for 1.5 tweezers&Lv1&\\\hline
  Home temp servo for 1.0 Innolume&Lv1&\\\hline
  Shutter driver&Lv1 on M2 controller&\\\hline
  TiSapph&Floor&\\\hline
  Function generator for 1.0 Na switching&Lv1 above TiSapph&Save setting\\\hline
  Keithley power for 1.0 Innolume&Lv2 above TiSapph&\\\hline
  Till driver v2 for 1.0 Cs MOT&Lv2 above 1.5 Cs MOT&\\\hline
  Home temp servo for 1.0 Cs MOT&Lv2 on Till driver v2&\\\hline
  Benchtop power&Lv2 facing computer&\\\hline
  High voltage power&Above Lv2 on power outlet&\\\hline
  \textbf{1.5 Rack}&&\\\hline
  Home power (24)&Bottom level&\\\hline
  SEAS pump controller&Middle left&\\\hline
  Benchtop power&Middle next to pump controller&\\\hline
  Oscilloscope for 1.5 monitor&Above pump controller&\\\hline
  Benchtop power for 1.5 coil&Above oscilloscope&\\\hline
  High voltage Amp for 1.5 MOT piezo&Top level&\\\hline
  Function generator for 1.5 MOT piezo&Top level on high voltage Amp&\\\hline
  \textbf{Below 1.5 table}&&\\\hline
  Benchtop power x3&Near outer edge&\\\hline
  IGBT stack&Floor&\\\hline
  LED driver&Near inner edge&\\\hline
  AWG computer&Floor&\\\hline
  \textbf{Above 1.5 table}&&\\\hline
  Benchtop power&Lv1&\\\hline
  Benchtop power&Lv2&\\\hline
  Radio computer&Lv2&\\\hline
\end{longtable}
\twocolumn

\subsection{Equipments}
\DevDef{TiSapph}{
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off pump, unplug from wall
  \item\OffBox Turn off controller, unplug from wall
  \item\OffBox Turn off chiller, unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into wall, turn on chiller
  \item\OnBox Plug into wall, turn on controller
  \item\OnBox Plug into wall, turn on pump
  \end{itemize}
}
\DevDef{Quantel Laser}{
  \textbf{1st level in the middle, facing 1.5 table}\\
  Ask Frederic.
  \vspace{1cm}
}
\DevDef{Intensity servo for Na}{
  \textbf{1st level above Na tweezer, below 1.0 computer control box}\\
  Unplug power
}
\DevDef{1st version Till's current controller for butterfly laser}{
  \textbf{1st level above TiSapph}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power (zip tie to controller)
  \end{itemize}
}
\DevDef{Homebuilt temperature servo for butterfly laser}{
  \textbf{1st level above TiSapph}\\
  Switch off/on the front switch
}
\DevDef{PLL for 1.5 Cs MOT}{
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug all signals (zip tie together)
  \item\OffBox Unplug 15V and 5V powers (zip tie together)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plugin 15V and 5V powers
  \item\OnBox Plugin all signals
  \item\OnBox Apply settings ???
  \end{itemize}
}
\subsection{Power supplies}
\DevDef{Homebuilt $24$V power supply}{
  \textbf{2nd level above NaCs 1.0 chamber, facing the drawers/wall}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Benchtop power supply for dispensers}{
  \textbf{1st level above NaCs 1.0 chamber}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Benchtop power supply for 1.5 computer control clock}{
  \textbf{2nd level next to 1.5 computer control box}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Homebuilt $\pm15,5$V power supply}{
  \textbf{2nd level above TiSapph}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}
\subsection{Computers}
\DevDef{Computer control box for 1.5}{
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox (Do following three steps quickly)
  \item\OffBox Turn off the box by flipping the switch on the front side of the box
  \item\OffBox Unplug the high voltage (48V) power supply (circular plug)
    on the left side of the box. Then unplug the power supply itself fromm the outlet.
  \item\OffBox Unplug the 12V board power supply.
    (L.T.E. one with a rectangular connector in the front)
  \item\OffBox Unplug USB power on the front side near the fan.
  \item\OffBox Zip tie the three unpluged powers to the computer control box.
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Connect to 3.5G Windfrek clock generator from Tweezer1.
    Set frequency to 3.5G and amplitude to max.
  \item\OnBox (Remove power connection zip ties and) Plug in USB power supply.
  \item\OnBox (Do following three steps quickly)
  \item\OnBox Plug in the 12V board power supply.
  \item\OnBox Plug in high voltage power supply. The fans should start spinning.
  \item\OnBox Turn on the power switch of the box.
  \end{itemize}
}

\end{document}