### 2019-04-26 presentations/2019-04-27_shutdown/checklist.tex
[presentation/2019-04-27_shutdown] WIP: checklists

\documentclass[10pt,fleqn,twocolumn]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{esint}
\usepackage{bbm}
\usepackage{amscd}
\usepackage{picinpar}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usepackage{indentfirst}
\usepackage{wrapfig}
\usepackage{units}
\usepackage{textcomp}
\usepackage[utf8x]{inputenc}
% \usepackage{feyn}
\usepackage{feynmp}
\usepackage{xkeyval}
\usepackage{xargs}
\usepackage{verbatim}
\usepackage{pgfplots}
\usepackage{hyperref}
\usepackage{longtable}

\usetikzlibrary{
  arrows,
  calc,
  decorations.pathmorphing,
  decorations.pathreplacing,
  decorations.markings,
  fadings,
  positioning,
  shapes
}

\DeclareGraphicsRule{*}{mps}{*}{}
\newcommand{\ud}{\mathrm{d}}
\newcommand{\ue}{\mathrm{e}}
\newcommand{\ui}{\mathrm{i}}
\newcommand{\res}{\mathrm{Res}}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\dsum}{\displaystyle\sum}
\newcommand{\dprod}{\displaystyle\prod}
\newcommand{\dlim}{\displaystyle\lim}
\newcommand{\dint}{\displaystyle\int}
\newcommand{\fsno}[1]{{\!\not\!{#1}}}
\newcommand{\eqar}[1]
{
  \begin{align*}
    #1
  \end{align*}
}
\newcommand{\texp}[2]{\ensuremath{{#1}\times10^{#2}}}
\newcommand{\dexp}[2]{\ensuremath{{#1}\cdot10^{#2}}}
\newcommand{\eval}[2]{{\left.{#1}\right|_{#2}}}
\newcommand{\paren}[1]{{\left({#1}\right)}}
\newcommand{\lparen}[1]{{\left({#1}\right.}}
\newcommand{\rparen}[1]{{\left.{#1}\right)}}
\newcommand{\abs}[1]{{\left|{#1}\right|}}
\newcommand{\sqr}[1]{{\left[{#1}\right]}}
\newcommand{\crly}[1]{{\left\{{#1}\right\}}}
\newcommand{\angl}[1]{{\left\langle{#1}\right\rangle}}
\newcommand{\tpdiff}[4][{}]{{\paren{\frac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}_{#4}}}
\newcommand{\tpsdiff}[4][{}]{{\paren{\frac{\partial^{#1}}{\partial {#3}{}^{#1}}{#2}}_{#4}}}
\newcommand{\pdiff}[3][{}]{{\frac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}}
\newcommand{\diff}[3][{}]{{\frac{\ud^{#1} {#2}}{\ud {#3}{}^{#1}}}}
\newcommand{\psdiff}[3][{}]{{\frac{\partial^{#1}}{\partial {#3}{}^{#1}} {#2}}}
\newcommand{\sdiff}[3][{}]{{\frac{\ud^{#1}}{\ud {#3}{}^{#1}} {#2}}}
\newcommand{\tpddiff}[4][{}]{{\left(\dfrac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}\right)_{#4}}}
\newcommand{\tpsddiff}[4][{}]{{\paren{\dfrac{\partial^{#1}}{\partial {#3}{}^{#1}}{#2}}_{#4}}}
\newcommand{\pddiff}[3][{}]{{\dfrac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}}
\newcommand{\ddiff}[3][{}]{{\dfrac{\ud^{#1} {#2}}{\ud {#3}{}^{#1}}}}
\newcommand{\psddiff}[3][{}]{{\frac{\partial^{#1}}{\partial{}^{#1} {#3}} {#2}}}
\newcommand{\sddiff}[3][{}]{{\frac{\ud^{#1}}{\ud {#3}{}^{#1}} {#2}}}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{fontenc}
\usepackage{titlesec}
% \usepackage{tipa}
\usepackage{ulem}
\usepackage{color}
\usepackage{cancel}

\newcommand{\hcancel}[2][black]{\setbox0=\hbox{#2}%
  \rlap{\raisebox{.45\ht0}{\textcolor{#1}{\rule{\wd0}{1pt}}}}#2}
\pagestyle{fancy}
% \setlength{\headheight}{60pt}
\fancyhead{}
\fancyfoot{}
\fancyfoot[C]{\thepage}
\fancyhead[R]{}
\renewcommand{\footruleskip}{0pt}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}
\titlespacing{\subsection}{0pt}{\parskip}{-\parskip}

\newcommand\pgfmathsinandcos[3]{%
  \pgfmathsetmacro#1{sin(#3)}%
  \pgfmathsetmacro#2{cos(#3)}%
}
\newcommand\LongitudePlane[3][current plane]{%
  \pgfmathsinandcos\sinEl\cosEl{#2} % elevation
  \pgfmathsinandcos\sint\cost{#3} % azimuth
  \tikzset{#1/.estyle={cm={\cost,\sint*\sinEl,0,\cosEl,(0,0)}}}
}
\newcommand\LatitudePlane[3][current plane]{%
  \pgfmathsinandcos\sinEl\cosEl{#2} % elevation
  \pgfmathsinandcos\sint\cost{#3} % latitude
  \pgfmathsetmacro\yshift{\cosEl*\sint}
  \tikzset{#1/.estyle={cm={\cost,0,0,\cost*\sinEl,(0,\yshift)}}} %
}
\newcommand\DrawLongitudeCircle[2][1]{
  \LongitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{atan(sin(#2)*cos(\angEl)/sin(\angEl))} %
  \draw[current plane] (\angVis:1) arc (\angVis:\angVis+180:1);
  \draw[current plane,dashed] (\angVis-180:1) arc (\angVis-180:\angVis:1);
}
\newcommand\DrawLatitudeCircleArrow[2][1]{
  \LatitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  \pgfmathsetmacro\sinVis{sin(#2)/cos(#2)*sin(\angEl)/cos(\angEl)}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{asin(min(1,max(\sinVis,-1)))}
  \draw[current plane,decoration={markings, mark=at position 0.6 with {\arrow{<}}},postaction={decorate},line width=.6mm] (\angVis:1) arc (\angVis:-\angVis-180:1);
  \draw[current plane,dashed,line width=.6mm] (180-\angVis:1) arc (180-\angVis:\angVis:1);
}
\newcommand\DrawLatitudeCircle[2][1]{
  \LatitudePlane{\angEl}{#2}
  \tikzset{current plane/.prefix style={scale=#1}}
  \pgfmathsetmacro\sinVis{sin(#2)/cos(#2)*sin(\angEl)/cos(\angEl)}
  % angle of "visibility"
  \pgfmathsetmacro\angVis{asin(min(1,max(\sinVis,-1)))}
  \draw[current plane] (\angVis:1) arc (\angVis:-\angVis-180:1);
  \draw[current plane,dashed] (180-\angVis:1) arc (180-\angVis:\angVis:1);
}
\newcommand\coil[1]{
  {\rh * cos(\t * pi r)}, {\apart * (2 * #1 + \t) + \rv * sin(\t * pi r)}
}
\makeatletter
\define@key{DrawFromCenter}{style}[{->}]{
  \tikzset{DrawFromCenterPlane/.style={#1}}
}
\define@key{DrawFromCenter}{r}[1]{
  \def\@R{#1}
}
\define@key{DrawFromCenter}{center}[(0, 0)]{
  \def\@Center{#1}
}
\define@key{DrawFromCenter}{theta}[0]{
  \def\@Theta{#1}
}
\define@key{DrawFromCenter}{phi}[0]{
  \def\@Phi{#1}
}
\presetkeys{DrawFromCenter}{style, r, center, theta, phi}{}
\newcommand*\DrawFromCenter[1][]{
  \setkeys{DrawFromCenter}{#1}{
    \pgfmathsinandcos\sint\cost{\@Theta}
    \pgfmathsinandcos\sinp\cosp{\@Phi}
    \pgfmathsinandcos\sinA\cosA{\angEl}
    \pgfmathsetmacro\DX{\@R*\cost*\cosp}
    \pgfmathsetmacro\DY{\@R*(\cost*\sinp*\sinA+\sint*\cosA)}
    \draw[DrawFromCenterPlane] \@Center -- ++(\DX, \DY);
  }
}
\newcommand*\DrawFromCenterText[2][]{
  \setkeys{DrawFromCenter}{#1}{
    \pgfmathsinandcos\sint\cost{\@Theta}
    \pgfmathsinandcos\sinp\cosp{\@Phi}
    \pgfmathsinandcos\sinA\cosA{\angEl}
    \pgfmathsetmacro\DX{\@R*\cost*\cosp}
    \pgfmathsetmacro\DY{\@R*(\cost*\sinp*\sinA+\sint*\cosA)}
    \draw[DrawFromCenterPlane] \@Center -- ++(\DX, \DY) node {#2};
  }
}
\makeatother
\tikzstyle{snakearrow} = [decorate, decoration={pre length=0.2cm,
  post length=0.2cm, snake, amplitude=.4mm,
  segment length=2mm},thick, ->]
%% document-wide tikz options and styles
\tikzset{%
  >=latex, % option for nice arrows
  inner sep=0pt,%
  outer sep=2pt,%
  mark coordinate/.style={inner sep=0pt,outer sep=0pt,minimum size=3pt,
    fill=black,circle}%
}
\addtolength{\hoffset}{-1.3cm}
\addtolength{\voffset}{-2cm}
\addtolength{\textwidth}{3cm}
\addtolength{\textheight}{2.5cm}
\renewcommand{\footskip}{10pt}
\setlength{\headwidth}{\textwidth}
\setlength{\headsep}{20pt}
\setlength{\marginparwidth}{0pt}

\makeatletter
\newcommand\ColorBox[2]{
  \begin{tikzpicture}
    \fill[#1] (0, 0) rectangle (\f@size/25, \f@size/25);
    \draw[color=gray!80] (0, 0) rectangle (\f@size/25, \f@size/25);
    \path (\f@size/50, \f@size/50) node[color=black!15] {\small #2};
  \end{tikzpicture}
}
\newcommand\DevBoxes{\mbox{{\Large\ColorBox{color=red!35}{Off}\!\!\ColorBox{color=green!35}{On}}}}
\newcommand\OffBox{{\normalsize\ColorBox{color=orange!40}{}}}
\newcommand\OnBox{{\normalsize\ColorBox{color=cyan!40}{}}}
\makeatother

\newcommand\DevHeader[1]{{\Large\DevBoxes\textbf{#1}\\}}
\def\arraystretch{2}

\newcommand\DevDef[2]{
  \begin{minipage}[t]{\linewidth}
    \begin{itemize}
    \item\DevHeader{#1}
      #2
    \end{itemize}
    \vspace{0.8cm}
  \end{minipage}
}
\newcommand\Person[2]{
\clearpage
\section*{#1}
\fancyhead[R]{#1}
\setcounter{page}{1}
\setcounter{subsection}{0}
For shutting off, go through the checklist in the following order
\begin{itemize}
\item\OffBox Turn off equipments
\item\OffBox Turn off power supplies
\item\OffBox Turn off computers
\end{itemize}
Wait before everyone else finish their corresponding steps before starting the next one.\\
For powering up, go through the checklist in the following order
\begin{itemize}
\item\OnBox Turn on computers
\item\OnBox Turn on power supplies
\item\OnBox Turn on equipments
\item\OnBox Check computer connection to devices
\end{itemize}
Also wait before everyone else finish their corresponding steps before starting the next one.\\

\textbf{\Large #2}
\vfill
}
\renewcommand{\thesubsection}{\arabic{subsection}}

\parindent=0pt
\title{Checklist}
\date{April 27th, 2019}
\begin{document}

\onecolumn
\Person{Yichao}{
  Equipments and power supplies are mostly located around the laser table facing the wall.
}
\begin{center}
  {\LARGE\textbf{Equipment list}}
\end{center}
\begin{longtable}{|c|c|c|}
  \hline
  \textbf{ID}&\textbf{Equipment}&\textbf{Location}\\\hline
  \multicolumn{3}{|c|}{\textbf{Laser table facing computer or the wall} west $\rightarrow$ south}\\\hline
  1&Benchtop power 2chn for Cs MOT beat lock&Lv1 above 1.0 Cs MOT\\\hline
  2&PLL for Cs MOT&Below Cs MOT\\\hline
  3&Home power ($\pm$15,5)&Lv2 above Cs MOT\\\hline
  4&Valan&1.5 amplifier stack Lv1 above 1.0 Cs MOT\\\hline
  5&Till driver v2 for 1.5 Na Raman&Lv1 above 1.5 Na Raman\\\hline
  6&Temp servo for 1.5 Na Raman seed&Lv1 above 1.5 Na Raman\\\hline
  7&Temp servo for 1.5 Na Raman doubler&Lv1 above 1.5 Na Raman\\\hline
  8&PLL for Cs Raman&Next to 1.0 Cs Raman\\\hline
  9&New Focus driver x2 for Cs Raman&Lv1 above 1.0 Cs Raman\\\hline
  10&Benchtop power (x3)&Lv2 above 1.0 Cs MOT\\\hline
  11&Home power (24V) x2&Lv2 above 1.0 Cs Raman\\\hline
  12&Benchtop power&Lv1 next to New Focus drivers\\\hline
  13&Shutter driver&Lv1 above 1.0 Cs Raman\\\hline
  14&New Focus lock box for 1.0 Cs Raman&Lv1 above 1.0 Cs Raman\\\hline
  15&Thorlabs temp servo for 1.0 Na Raman doubler&Lv1 above 1.0 Na seed\\\hline
  16&Timebase drivers for 1.0 Na Raman and Na D2&Lv1 above Thorlabs temp servo\\\hline
  17&Benchtop power for 1.0 Cs Raman&Lv2 above New Focus lock box\\\hline
  18&High voltage power&Lv2 above timebase drivers\\\hline
  19&Home lock box for Na D2&Lv1 above 1.0 Na seed\\\hline
  20&MPB Raman laser&Lv1 above 1.0 Na MOT\\\hline
  21&Oscilloscope for Na D2 lock&Lv1 above MPB Raman laser\\\hline
  22&Benchtop power for 1.0 Na Raman seed&Lv2 above oscilloscope\\\hline
\end{longtable}
\twocolumn

\subsection{Equipments}
\DevDef{PLL for Cs MOT [2]}{
  \textbf{Below Cs MOT}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug all signals (zip tie together)
  \item\OffBox Unplug 15V and 5V powers (zip tie together)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plugin 15V and 5V powers
  \item\OnBox Plugin all signals
  \item\OnBox Apply settings
  \end{itemize}
}
\DevDef{Valan [4]}{
  \textbf{1.5 amplifier stack Lv1 above 1.0 Cs MOT}\\
  Check with 1.5 to record setting\\

  \begin{tabular}{|c|c|}
    \hline
    Frequency&\hphantom{\hspace{3.5cm}}\\\hline
    Amplitude&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Restore settings
  \end{itemize}
}
\DevDef{Till driver v2 for 1.5 Na Raman [5]}{
  \textbf{Lv1 above 1.5 Na Raman}\\
  Record current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Current&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power (zip tie to controller)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Turn on
  \end{itemize}
}
\DevDef{Temp servo for 1.5 Na Raman seed [6]}{
  \textbf{Lv1 above 1.5 Na Raman}\\
  Check plugged into UPS.
}
\DevDef{Temp servo for 1.5 Na Raman doubler [7]}{
  \textbf{Lv1 above 1.5 Na Raman}\\
  Check plugged into UPS.
}
\DevDef{PLL for Cs Raman [8]}{
  \textbf{Next to 1.0 Cs Raman}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug all signals (zip tie together)
  \item\OffBox Unplug 15V and 5V powers (zip tie together)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plugin 15V and 5V powers
  \item\OnBox Plugin all signals
  \item\OnBox Apply settings
  \end{itemize}
}
\DevDef{New Focus driver x2 for Cs Raman [9]}{
  \textbf{Lv1 above 1.0 Cs Raman}\\
  Record Temperature, Current, Piezo voltage\\
  Temperature is accessible in the menu under system status\\

  \begin{tabular}{|c|c|}
    \hline
    F3 Current&\hphantom{\hspace{3.5cm}}\\\hline
    F3 Piezo Voltage&\\\hline
    F3 Temperature&\\\hline
    F4 Current&\\\hline
    F4 Piezo Voltage&\\\hline
    F4 Temperature&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off output.
  \item\OffBox Check plugged into UPS.
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox (Turn on)
  \end{itemize}
}
\DevDef{Shutter driver [13]}{
  \textbf{Lv1 above 1.0 Cs Raman}\\
  Switch off/on with the switch on the back.
}
\DevDef{New Focus lock box for 1.0 Cs Raman [14]}{
  \textbf{Lv1 above 1.0 Cs Raman}\\
  Unplug power.
}
\DevDef{Thorlabs temp servo for 1.0 Na Raman doubler [15]}{
  \textbf{Lv1 above 1.0 Na seed}\\
  Record temperature settings\\

  \begin{tabular}{|c|c|}
    \hline
    Setpoint&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  Check plugged into UPS.
}

\DevDef{Timebase drivers for 1.0 Na Raman and Na D2 [16]}{
  \textbf{Lv1 above Thorlabs temp servo}\\
  Record current and T1 settings\\

  \begin{tabular}{|c|c|}
    \hline
    Raman Current&\hphantom{\hspace{3.5cm}}\\\hline
    Raman T1&\\\hline
    D2 Current&\hphantom{\hspace{3.5cm}}\\\hline
    D2 T1&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off output.
  \item\OffBox Check plugged into UPS.
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Turn on
  \end{itemize}
}

\DevDef{Home lock box for Na D2 [19]}{
  \textbf{Lv1 above 1.0 Na seed}\\
  \textbf{Turn on-off together with the adjustable high voltage power supply above}\\
  Record high voltage power supply output

  \begin{tabular}{|c|c|}
    \hline
    Voltage&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unlock
  \item\OffBox Turn down the voltage of the high voltage power supply to 0
  \item\OffBox Unplug high voltage power on lockbox (zip tie)
  \item\OffBox Unplug 15V from lockbox (zip tie)
  \item\OffBox Turn off high voltage power supply
  \item\OffBox Unplug high voltage power supply from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug high voltage power supply into the wall
  \item\OnBox Turn on high voltage power supply (make sure the output is 0)
  \item\OnBox Plug in 15V to lockbox
  \item\OnBox Plug in high voltage to lockbox
  \item\OnBox Turn up the voltage of the high voltage power supply
  \item\OnBox Check HV output
  \end{itemize}
}

\DevDef{MPB Raman laser [20]}{
  \textbf{Lv1 above 1.0 Na MOT}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Power off
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into wall
  \item\OnBox Power on
  \end{itemize}
}
\DevDef{Oscilloscope for Na D2 lock [21]}{
  \textbf{Lv1 above MPB Raman laser}\\

  \begin{itemize}
  \item\OffBox Unplug from wall
  \end{itemize}
}
\subsection{Power supplies}
\DevDef{Benchtop power 2chn for Cs MOT beat lock [1]}{
  \textbf{Lv1 above 1.0 Cs MOT}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Home power ($\pm$15,5) [3]}{
  \textbf{Lv2 above Cs MOT}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Benchtop power (x3) [10]}{
  \textbf{Lv2 above 1.0 Cs MOT}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    No. 1 Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    No. 1 Current&\\\hline
    No. 2 Voltage&\\\hline
    No. 2 Current&\\\hline
    No. 3 Voltage&\\\hline
    No. 3 Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Home power (24V) x2 [11]}{
  \textbf{Lv2 above 1.0 Cs Raman}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Benchtop power [12]}{
  \textbf{Lv1 next to New Focus driver}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Benchtop power for 1.0 Cs Raman [17]}{
  \textbf{Lv2 above New Focus lock box}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{High voltage power [18]}{
  \textbf{Lv2 above timebase drivers}\\
  Follow home lock box for Na D2 [19].
}

\DevDef{Benchtop power for 1.0 Na Raman seed [22]}{
  \textbf{Lv2 above oscilloscope}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Current&\\\hline
  \end{tabular}\\

  Check plugged into UPS.
}
\subsection{Computers}
\DevDef{Control computers}{
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Check update.
  \item\OffBox Turn off.
  \item\OffBox Unplug from UPS.
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Get UPS back
  \item\OnBox Turn on.
  \end{itemize}
}
\DevDef{Computer UPS}{
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Move to the lab.
  \item\OffBox Plug into backup power.
  \item\OffBox Plug in pump/temp controllers.
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug load back into the wall (or where they belongs).
  \item\OnBox Move back.
  \item\OnBox Plug computers back in.
  \end{itemize}
}

\onecolumn
\Person{Kenneth}{
  Equipments and power supplies are mostly located between the machine table and the laser table.
}
\begin{center}
  {\LARGE\textbf{Equipment list}}
\end{center}
\begin{longtable}{|c|c|c|}
  \hline
  \textbf{ID}&\textbf{Equipment}&\textbf{Location}\\\hline
  \multicolumn{3}{|c|}{\textbf{Laser table side facing machine table} north $\rightarrow$ east}\\\hline
  1&671 EOM stack&Floor\\\hline
  2&1550 Amp&Floor\\\hline
  3&Radio computer&Lv2\\\hline
  4&Benchtop power 2chn&Lv2 next to radio computer\\\hline
  5&High voltage power&Lv2 next to benchtop power\\\hline
  6&Variac for Na cell&Lv1 above 1.0 Na MOT\\\hline
  7&Shutter driver&Lv1 next to variac\\\hline
  8&Home lock box for 1.0 Na D1&Lv1 next to shutter driver\\\hline
  9&1038 Amp&Floor\\\hline
  10&1038 Timebase laser&Table\\\hline
  11&FPGA box for 1.0&Lv2\\\hline
  12&Thorlabs temp servo for Na D1 doubler&Lv1 above Na D1\\\hline
  13&Thorlabs temp servo for Na D1 seed&Lv1 above Na D1\\\hline
  14&Thorlabs current for Na D1 seed&Lv1 above Na D1\\\hline
  15&Benchtop power 2chn&Lv2 above Thorlabs controllers for Na D1 seed\\\hline
  16&Home power (24)&Lv2 next to benchtop power\\\hline
  17&Benchtop power x2 (2chn x1)&Lv2 next to Home power (24)\\\hline
  18&Till driver v2 for 976&Lv1 above Na D1 doubler\\\hline
  19&Thorlabs temp servo for Cs tweezer&Lv1 above Timebase 1038 beam\\\hline
  20&Thorlabs current (2A) (off)&Lv1 above Thorlabs temp servo\\\hline
  21&Till driver v2 for 671&Lv1 above Thorlabs current\\\hline
  22&Temp servo x2&Lv1 above Timebase 1038 control box\\\hline
  23&Oscilloscope&Lv1\\\hline
  \multicolumn{3}{|c|}{\textbf{Machine table side facing laser table} south}\\\hline
  24&Thorlabs current for 1.5 Cs MOT&Lv1 above 1.5 Cs\\\hline
  25&Till driver v2 for 1.5 Cs RP&Lv1 above 1.5 Cs\\\hline
  26&Temp servo (x2) for 1.5 Cs MOT/RP&Lv1 above 1.5 Cs\\\hline
  27&Oscilloscope for 1.5 Cs&Lv1 above 1.5 Cs\\\hline
  28&Home lock box for 1.5 Cs&Lv1 above 1.5 Cs\\\hline
  29&Novatec for 1.5 Cs lock&Lv1 above 1.5 Cs\\\hline
  30&Benchtop power 2chn&Lv2 1.5 Cs MOT/RP current\\\hline
  31&Home power (24) x2&Lv2 next to benchtop power\\\hline
  32&Benchtop power&Lv2 on two home power (24)\\\hline
  33&Home power ($\pm$15,5)&Lv2 middle\\\hline
  34&Benchtop power (x2) for uWave Amp&Lv2 next to home power ($\pm$15,5)\\\hline
  35&Benchtop power (x2)&Lv1 above 1.5 Na\\\hline
  36&High voltage power&Lv1 next to 2 benchtop power\\\hline
  37&Function generator for 1.0 MOT piezo&Lv1\\\hline
  38&Temperature and pump controllers for cavity&Lv1 next to function generator\\\hline
  39&Water chiller for uWave breadboard&Floor\\\hline
  40&SAES pump&Lv1 above 1.0 chamber\\\hline
  41&Benchtop power x2 (2chn x1)&Lv2 next to coil servo\\\hline
\end{longtable}
\twocolumn
\subsection{Equipments}
\DevDef{671 EOM stack [1]}{
  \textbf{Floor}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    No. 1 Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    No. 1 Left Current&\\\hline
    No. 1 Right Voltage&\\\hline
    No. 1 Right Current&\\\hline
    Thorlabs Current&\\\hline
    No. 2 Voltage&\\\hline
    No. 2 Current&\\\hline
    &\\\hline
    &\\\hline
    &\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \item\OffBox
  \item\OffBox
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \item\OnBox
  \item\OnBox
  \end{itemize}
}

\DevDef{1550 Amp [2]}{
  \textbf{Floor}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox
  \item\OffBox
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox
  \item\OnBox
  \end{itemize}
}

\DevDef{Variac for Na cell [6]}{
  \textbf{Lv1 above 1.0 Na MOT}\\
  Record temperature and voltage settings\\

  \begin{tabular}{|c|c|}
    \hline
    Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Temperature&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  Check if we can use the backup power. (Through the UPS)
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn voltage to 0
  \item\OffBox Switch off
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into wall
  \item\OnBox Switch on
  \item\OnBox Turn the voltage up \textbf{SLOWLY} and monitor the temperature at the same time.
  \end{itemize}
}

\DevDef{Shutter driver [7]}{
  \textbf{Lv1 next to variac}\\
  Switch off/on with the switch on the back.
}

\DevDef{Home lock box for 1.0 Na D1 [8]}{
  \textbf{Lv1 next to shutter driver}\\
  \textbf{Turn on-off together with the fixed high voltage power supply above}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unlock
  \item\OffBox Unplug high voltage power on lockbox (zip tie)
  \item\OffBox Unplug 15V from lockbox (zip tie)
  \item\OffBox Unplug high voltage power supply from walls
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug high voltage power supply into the wall
  \item\OnBox Plug in 15V to lockbox
  \item\OnBox Plug in high voltage to lockbox
  \item\OnBox Check HV output
  \end{itemize}
}

\DevDef{1038 Amp [9]}{
  \textbf{Floor}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox
  \item\OffBox
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox
  \item\OnBox
  \end{itemize}
}

\DevDef{1038 Timebase laser [10]}{
  \textbf{Table}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox
  \item\OffBox
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox
  \item\OnBox
  \end{itemize}
}

\DevDef{Thorlabs temp servo for Na D1 doubler [12]}{
  \textbf{Lv1 above Na D1}\\
  Record temperature settings\\

  \begin{tabular}{|c|c|}
    \hline
    Setpoint&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  Check plugged into UPS.
}

\DevDef{Thorlabs temp servo for Na D1 seed [13]}{
  \textbf{Lv1 above Na D1}\\
  Record temperature settings\\

  \begin{tabular}{|c|c|}
    \hline
    Setpoint&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  Check plugged into UPS.
}

\DevDef{Thorlabs current for Na D1 seed [14]}{
  \textbf{Lv1 above Na D1}\\
  Record current settings\\

  \begin{tabular}{|c|c|}
    \hline
    Current&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Check current setting
  \item\OnBox Turn on
  \end{itemize}
}

\DevDef{Till driver v2 for 976 [18]}{
  \textbf{Lv1 above Na D1 doubler}\\
  Record current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Current&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power (zip tie to controller)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Turn on
  \end{itemize}
}

\DevDef{Thorlabs temp servo for Cs tweezer [19]}{
  \textbf{Lv1 above Timebase 1038 beam}\\
  Record temperature settings\\

  \begin{tabular}{|c|c|}
    \hline
    Setpoint&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  Check plugged into UPS.
}

\DevDef{Till driver v2 for 671 [21]}{
  \textbf{Lv1 above Thorlabs current}\\
  Record current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Current&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power (zip tie to controller)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Turn on
  \end{itemize}
}

\DevDef{Temp servo x2 [22]}{
  \textbf{Lv1 above Timebase 1038 control box}\\
  Check plugged into UPS.
}

\DevDef{Oscilloscope [23]}{
  \textbf{Lv1}\\

  \begin{itemize}
  \item\OffBox Unplug from wall
  \end{itemize}
}

\DevDef{Thorlabs current for 1.5 Cs MOT [24]}{
  \textbf{Lv1 above 1.5 Cs}\\
  Record current settings\\

  \begin{tabular}{|c|c|}
    \hline
    Current&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Check current setting
  \item\OnBox Turn on
  \end{itemize}
}

\DevDef{Till driver v2 for 1.5 Cs RP [25]}{
  \textbf{Lv1 above 1.5 Cs}\\
  Record current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Current&\hphantom{\hspace{3.5cm}}\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power (zip tie to controller)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in power
  \item\OnBox Turn on
  \end{itemize}
}

\DevDef{Temp servo (x2) for 1.5 Cs MOT/RP [26]}{
  \textbf{Lv1 above 1.5 Cs}\\
  Check plugged into UPS.
}

\DevDef{Oscilloscope for 1.5 Cs [27]}{
  \textbf{Lv1 above 1.5 Cs}\\

  \begin{itemize}
  \item\OffBox Unplug from wall
  \end{itemize}
}

\DevDef{Home lock box for 1.5 Cs [28]}{
  \textbf{Lv1 above 1.5 Cs}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unlock
  \item\OffBox Unplug 15V from lockbox (zip tie)
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug in 15V to lockbox
  \end{itemize}
}

\DevDef{Novatec for 1.5 Cs lock [29]}{
  \textbf{Lv1 above 1.5 Cs}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Save settings
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Restore settings
  \end{itemize}
}

\DevDef{Function generator for 1.0 MOT piezo [37]}{
  \textbf{Lv1}\\

  \textbf{Turn on-off together with the HV amplifiers for 1.0 piezo mirrors above
    and the high voltage power supply on the left}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off the function generator
  \item\OffBox Unplug function generator output (zip tie and label)
  \item\OffBox Unplug HV power supply from the wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plugin HV power supply to the wall
  \item\OnBox Plugin function generator output
  \item\OnBox Turn on the function generator
  \end{itemize}
}

\DevDef{Temperature and pump controllers for cavity [38]}{
  \textbf{Lv1 next to function generator}\\

  Check plugged into UPS.
}

\DevDef{Water chiller for uWave breadboard [39]}{
  \textbf{Floor}\\
}

\DevDef{SAES pump [40]}{
  \textbf{Lv1 above 1.0 chamber}\\

  Check plugged into UPS.
}
\subsection{Power supplies}
\DevDef{Benchtop power 2chn [4]}{
  \textbf{Lv2 next to radio computer}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{High voltage power [5]}{
  \textbf{Lv2 next to benchtop power}\\
  Follow home lock box for Na D2 [8].
}

\DevDef{Benchtop power 2chn [15]}{
  \textbf{Lv2 above Thorlabs controllers for Na D1 seed}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Home power (24) [16]}{
  \textbf{Lv2 next to benchtop power}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Benchtop power x2 (2chn x1) [17]}{
  \textbf{Lv2 next to Home power (24)}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    No. 1 Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    No. 1 Current&\\\hline
    No. 2 Left Voltage&\\\hline
    No. 2 Left Current&\\\hline
    No. 2 Right Voltage&\\\hline
    No. 2 Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Benchtop power 2chn [30]}{
  \textbf{Lv2 1.5 Cs MOT/RP current}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Left Current&\\\hline
    Right Voltage&\\\hline
    Right Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Home power (24) x2 [31]}{
  \textbf{Lv2 next to benchtop power}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Benchtop power [32]}{
  \textbf{Lv2 on two home power (24)}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Home power ($\pm$15,5) [33]}{
  \textbf{Lv2 middle}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{Benchtop power (x2) for uWave Amp [34]}{
  \textbf{Lv2 next to home power ($\pm$15,5)}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    No. 1 Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    No. 1 Current&\\\hline
    No. 2 Voltage&\\\hline
    No. 2 Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug uWave input (tie).
  \item\OffBox Turn low voltage one to (-)5V.
  \item\OffBox Turn down higher voltage one to 0.
  \item\OffBox Unplug load, label/tie.
  \item\OffBox Switch off power supplies.
  \item\OffBox Unplug from wall.
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall.
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}.
  \item\OnBox Plug -5V.
  \item\OnBox Plug higher voltage one.
  \item\OnBox Turn higher voltage up to original value.
  \item\OnBox Turn low voltage one down.
  \item\OnBox Plug uWave back in.
  \end{itemize}
}

\DevDef{Benchtop power (x2) [35]}{
  \textbf{Lv1 above 1.5 Na}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    No. 1 Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    No. 1 Current&\\\hline
    No. 2 Voltage&\\\hline
    No. 2 Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}

\DevDef{High voltage power [36]}{
  \textbf{Lv1 next to 2 benchtop power}
}

\DevDef{Benchtop power x2 (2chn x1) [41]}{
  \textbf{Lv2 next to coil servo}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    No. 1 Left Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    No. 1 Left Current&\\\hline
    No. 1 Right Voltage&\\\hline
    No. 1 Right Current&\\\hline
    No. 2 Voltage&\\\hline
    No. 2 Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\subsection{Computers}
\DevDef{Radio computer [3]}{
  \textbf{Lv2}\\
}

\DevDef{FPGA box for 1.0 [11]}{
  \textbf{Lv2}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox (Do following three steps quickly)
  \item\OffBox Turn off the box by flipping the switch on the front side of the box
  \item\OffBox Unplug the high voltage (48V) power supply (circular plug)
    on the front side of the box. Then unplug the power supply itself fromm the outlet.
  \item\OffBox Unplug the 12V board power supply.
    (L.T.E. one with a rectangular connector in the front)
  \item\OffBox Unplug USB power on the front side near the fan.
  \item\OffBox Zip tie the three unpluged powers to the computer control box.
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Connect to 3.5G Windfrek clock generator from Tweezer1.
    Set frequency to 3.5G and amplitude to max.
  \item\OnBox (Remove power connection zip ties and) Plug in USB power supply.
  \item\OnBox (Do following three steps quickly)
  \item\OnBox Plug in the 12V board power supply.
  \item\OnBox Plug in high voltage power supply.
  \item\OnBox Turn on the power switch of the box. The fans should start spinning.
  \end{itemize}
}

\onecolumn
\Person{Jessie}{
  Equipments and power supplies are mostly located between the machine table and the 1.5 table.
}
\begin{center}
  {\LARGE\textbf{Equipment list}}
\end{center}
\begin{longtable}{|c|c|c|}
  \hline
  \textbf{ID}&\textbf{Equipment}&\textbf{Location}\\\hline
  \multicolumn{3}{|c|}{\textbf{Machine table Facing the wall} north}\\\hline
  1&Benchtop power 2chn&Lv2\\\hline
  2&Home power (24)&Lv2 above 1.0 tweezer breadboard\\\hline
  3&Home power ($\pm$15,5)&Lv2 above home power (24)\\\hline
  4&Benchtop power x9 (2chn x3)&Lv2 next to home power (24) and ($\pm$15,5)\\\hline
  5&Current servo for compensation coil&Lv1 below home power (24)\\\hline
  6&Kenneth servo for 1.0 Cs tweezer&Lv1\\\hline
  7&High voltage Amp for 630 lock&Lv1\\\hline
  8&Vincent servo x2 for STIRAP lasers&Lv1\\\hline
  9&Greiner servo for 1.0 Na tweezer&Lv1 above Vincent servos\\\hline
  10&Rack for STRAP&Lv1 next to Vincent servos\\\hline
  11&Till driver v2 x2 for STIRAP lasers&Lv1 above rack for 1.5\\\hline
  12&Temp servo x3 for STIRAP&Lv2 next to 9x benchtop power\\\hline
  13&FPGA box for 1.5&Lv2\\\hline
  14&Home power ($\pm$15,5)&Lv2 next to FPGA box\\\hline
  15&Benchtop power x2&Lv2 on home power ($\pm$15,5)\\\hline
  16&Benchtop power&Lv2 next to home power ($\pm$15,5)\\\hline
  17&Benchtop power x2 (2chn x1)&Floor rack Lv1\\\hline
  18&Till driver v2 x2 for 1.5 Cs Raman&Floor rack Lv2\\\hline
  19&Kenneth servo for 1.5 Cs Raman phase lock&Floor rack Lv2 above Till drivers\\\hline
  20&Benchtop power&Floor rack Lv2 next to Till drivers\\\hline
  21&Temp servo x2 for 1.5 Cs Raman&Floor rack Lv2\\\hline
  22&Piezo driver for 1.5 Cs Raman&Floor rack Lv2 above home temp servos\\\hline
  23&Delay generator for 1.5 Na&Lv1 below 1.5 output panel\\\hline
  24&Till driver v2 for 1.0 Innolume&Lv1 next to delay generator\\\hline
  25&Kenneth servo for 1.5 tweezers&Lv1\\\hline
  26&Temp servo for 1.0 Innolume&Lv1\\\hline
  27&Shutter driver&Lv1 on M2 controller\\\hline
  28&TiSapph&Floor\\\hline
  29&Function generator for 1.0 Na switching&Lv1 above TiSapph\\\hline
  30&Keithley power for 1.0 Innolume&Lv2 above TiSapph\\\hline
  31&Till driver v2 for 1.0 Cs MOT&Lv2 above 1.5 Cs MOT\\\hline
  32&Temp servo for 1.0 Cs MOT&Lv2 on Till driver v2\\\hline
  33&Benchtop power 2chn&Lv2 facing computer\\\hline
  34&High voltage power&Above Lv2 on power outlet\\\hline
  \multicolumn{3}{|c|}{1.5 Rack}\\\hline
  35&Home power (24)&Bottom level\\\hline
  36&SEAS pump controller&Middle left\\\hline
  37&Benchtop power&Middle next to pump controller\\\hline
  38&Oscilloscope for 1.5 monitor&Above pump controller\\\hline
  39&Benchtop power 2chn for 1.5 coil&Above oscilloscope\\\hline
  40&High voltage Amp for 1.5 MOT piezo&Top level\\\hline
  41&Function generator for 1.5 MOT piezo&Top level on high voltage Amp\\\hline
  \multicolumn{3}{|c|}{Below 1.5 table}\\\hline
  42&Benchtop power x3&Near outer edge\\\hline
  43&IGBT stack&Floor\\\hline
  44&LED driver&Near inner edge\\\hline
  45&AWG computer&Floor\\\hline
  \multicolumn{3}{|c|}{Above 1.5 table}\\\hline
  46&Benchtop power&Lv1\\\hline
  47&Benchtop power&Lv2\\\hline
  48&Radio computer&Lv2\\\hline
\end{longtable}
\twocolumn

\subsection{Equipments}
\DevDef{Current servo for compensation coil [5]}{
  \textbf{Lv1 below home power (24)}\\
}

\DevDef{Kenneth servo for 1.0 Cs tweezer [6]}{
  \textbf{Lv1}\\
}

\DevDef{High voltage Amp for 630 lock [7]}{
  \textbf{Lv1}\\
}

\DevDef{Vincent servo x2 for STIRAP lasers [8]}{
  \textbf{Lv1}\\
}

\DevDef{Greiner servo for 1.0 Na tweezer [9]}{
  \textbf{Lv1 above Vincent servos}\\
}

\DevDef{Rack for STRAP [10]}{
  \textbf{Lv1 next to Vincent servos}\\
}

\DevDef{Till driver v2 x2 for STIRAP lasers [11]}{
  \textbf{Lv1 above rack for 1.5}\\
}

\DevDef{Temp servo x3 for STIRAP [12]}{
  \textbf{Lv2 next to 9x benchtop power}\\
}

\DevDef{FPGA box for 1.5 [13]}{
  \textbf{Lv2}\\
}

\DevDef{Till driver v2 x2 for 1.5 Cs Raman [18]}{
  \textbf{Floor rack Lv2}\\
}

\DevDef{Kenneth servo for 1.5 Cs Raman phase lock [19]}{
  \textbf{Floor rack Lv2 above Till drivers}\\
}

\DevDef{Temp servo x2 for 1.5 Cs Raman [21]}{
  \textbf{Floor rack Lv2}\\
}

\DevDef{Piezo driver for 1.5 Cs Raman [22]}{
  \textbf{Floor rack Lv2 above home temp servos}\\
}

\DevDef{Delay generator for 1.5 Na [23]}{
  \textbf{Lv1 below 1.5 output panel}\\
}

\DevDef{Till driver v2 for 1.0 Innolume [24]}{
  \textbf{Lv1 next to delay generator}\\
}

\DevDef{Kenneth servo for 1.5 tweezers [25]}{
  \textbf{Lv1}\\
}

\DevDef{Temp servo for 1.0 Innolume [26]}{
  \textbf{Lv1}\\
}

\DevDef{Shutter driver [27]}{
  \textbf{Lv1 on M2 controller}\\
}

\DevDef{TiSapph [28]}{
  \textbf{Floor}\\
}

\DevDef{Function generator for 1.0 Na switching [29]}{
  \textbf{Lv1 above TiSapph}\\
}

\DevDef{Keithley power for 1.0 Innolume [30]}{
  \textbf{Lv2 above TiSapph}\\
}

\DevDef{Till driver v2 for 1.0 Cs MOT [31]}{
  \textbf{Lv2 above 1.5 Cs MOT}\\
}

\DevDef{Temp servo for 1.0 Cs MOT [32]}{
  \textbf{Lv2 on Till driver v2}\\
}

\DevDef{SEAS pump controller [36]}{
  \textbf{Middle left}\\
}

\DevDef{Oscilloscope for 1.5 monitor [38]}{
  \textbf{Above pump controller}\\
}

\DevDef{High voltage Amp for 1.5 MOT piezo [40]}{
  \textbf{Top level}\\
}

\DevDef{Function generator for 1.5 MOT piezo [41]}{
  \textbf{Top level on high voltage Amp}\\
}

\DevDef{IGBT stack [43]}{
  \textbf{Floor}\\
}

\DevDef{LED driver [44]}{
  \textbf{Near inner edge}\\
}

\DevDef{TiSapph}{
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off pump, unplug from wall
  \item\OffBox Turn off controller, unplug from wall
  \item\OffBox Turn off chiller, unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into wall, turn on chiller
  \item\OnBox Plug into wall, turn on controller
  \item\OnBox Plug into wall, turn on pump
  \end{itemize}
}
\DevDef{Quantel Laser}{
  \textbf{1st level in the middle, facing 1.5 table}\\
  Ask Frederic.
  \vspace{1cm}
}
\DevDef{Intensity servo for Na}{
  \textbf{1st level above Na tweezer, below 1.0 computer control box}\\
  Unplug power
}
\DevDef{1st version Till's current controller for butterfly laser}{
  \textbf{1st level above TiSapph}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Turn off
  \item\OffBox Unplug power (zip tie to controller)
  \end{itemize}
}
\DevDef{Homebuilt temperature servo for butterfly laser}{
  \textbf{1st level above TiSapph}\\
  Switch off/on the front switch
}
\subsection{Power supplies}
\DevDef{Benchtop power 2chn [1]}{
  \textbf{Lv2}\\
}

\DevDef{Home power (24) [2]}{
  \textbf{Lv2 above 1.0 tweezer breadboard}\\
}

\DevDef{Home power ($\pm$15,5) [3]}{
  \textbf{Lv2 above home power (24)}\\
}

\DevDef{Benchtop power x9 (2chn x3) [4]}{
  \textbf{Lv2 next to home power (24) and ($\pm$15,5)}\\
}

\DevDef{Home power ($\pm$15,5) [14]}{
  \textbf{Lv2 next to FPGA box}\\
}

\DevDef{Benchtop power x2 [15]}{
  \textbf{Lv2 on home power ($\pm$15,5)}\\
}

\DevDef{Benchtop power [16]}{
  \textbf{Lv2 next to home power ($\pm$15,5)}\\
}

\DevDef{Benchtop power x2 (2chn x1) [17]}{
  \textbf{Floor rack Lv1}\\
}

\DevDef{Benchtop power [20]}{
  \textbf{Floor rack Lv2 next to Till drivers}\\
}

\DevDef{Benchtop power 2chn [33]}{
  \textbf{Lv2 facing computer}\\
}

\DevDef{High voltage power [34]}{
  \textbf{Above Lv2 on power outlet}\\
}

\DevDef{Home power (24) [35]}{
  \textbf{Bottom level}\\
}

\DevDef{Benchtop power [37]}{
  \textbf{Middle next to pump controller}\\
}

\DevDef{Benchtop power 2chn for 1.5 coil [39]}{
  \textbf{Above oscilloscope}\\
}

\DevDef{Benchtop power x3 [42]}{
  \textbf{Near outer edge}\\
}

\DevDef{Benchtop power [46]}{
  \textbf{Lv1}\\
}

\DevDef{Benchtop power [47]}{
  \textbf{Lv2}\\
}

\DevDef{Homebuilt $24$V power supply}{
  \textbf{2nd level above NaCs 1.0 chamber, facing the drawers/wall}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Benchtop power supply for dispensers}{
  \textbf{1st level above NaCs 1.0 chamber}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Benchtop power supply for 1.5 computer control clock}{
  \textbf{2nd level next to 1.5 computer control box}\\
  Record voltage/current values:\\

  \begin{tabular}{|c|c|}
    \hline
    Voltage&\hphantom{\hspace{3.5cm}}\\\hline
    Current&\\\hline
  \end{tabular}\\

  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug loads
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Turn off the power supply
  \item\OffBox Unplug from wall
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Plug into the wall
  \item\OnBox Turn on the power supply \textbf{Do NOT do this with loads pluged in}
  \item\OnBox Plug the load back in
  \end{itemize}
}
\DevDef{Homebuilt $\pm15,5$V power supply}{
  \textbf{2nd level above TiSapph}\\
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox Unplug load
  \item\OffBox Label and zip tie loads to the power supply
  \item\OffBox Switch off on the back
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Switch on
  \item\OnBox Plug the load back in
  \end{itemize}
}
\subsection{Computers}
\DevDef{AWG computer [45]}{
  \textbf{Floor}\\
}

\DevDef{Radio computer [48]}{
  \textbf{Lv2}\\
}

\DevDef{Computer control box for 1.5}{
  \textbf{Turn off check list}
  \begin{itemize}
  \item\OffBox (Do following three steps quickly)
  \item\OffBox Turn off the box by flipping the switch on the front side of the box
  \item\OffBox Unplug the high voltage (48V) power supply (circular plug)
    on the left side of the box. Then unplug the power supply itself fromm the outlet.
  \item\OffBox Unplug the 12V board power supply.
    (L.T.E. one with a rectangular connector in the front)
  \item\OffBox Unplug USB power on the front side near the fan.
  \item\OffBox Zip tie the three unpluged powers to the computer control box.
  \end{itemize}
  \textbf{Turn on check list}
  \begin{itemize}
  \item\OnBox Connect to 3.5G Windfrek clock generator from Tweezer1.
    Set frequency to 3.5G and amplitude to max.
  \item\OnBox (Remove power connection zip ties and) Plug in USB power supply.
  \item\OnBox (Do following three steps quickly)
  \item\OnBox Plug in the 12V board power supply.
  \item\OnBox Plug in high voltage power supply. The fans should start spinning.
  \item\OnBox Turn on the power switch of the box.
  \end{itemize}
}

\end{document}